# Zebra mussel network modeling script

# Contact information

Contact: Steven Brownlee
Email: steven.fr.brownlee@gmail.com
Date last revised: June 24 2024

# Session info

R version 4.4.1 (2024-06-14)
Platform: x86_64-redhat-linux-gnu
Running under: Nobara Linux 40 (GNOME Edition)

Matrix products: default
BLAS/LAPACK: FlexiBLAS OPENBLAS-OPENMP;  LAPACK version 3.11.0

locale:
 [1] LC_CTYPE=en_CA.UTF-8      LC_NUMERIC=C              LC_TIME=en_CA.utf8       
 [4] LC_COLLATE=en_CA.UTF-8    LC_MONETARY=en_CA.utf8    LC_MESSAGES=en_CA.UTF-8  
 [7] LC_PAPER=en_CA.utf8       LC_NAME=C                 LC_ADDRESS=C             
[10] LC_TELEPHONE=C            LC_MEASUREMENT=en_CA.utf8 LC_IDENTIFICATION=C      

time zone: America/Vancouver
tzcode source: system (glibc)

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

loaded via a namespace (and not attached):
 [1] digest_0.6.36     tidyr_1.3.1       utf8_1.2.4        R6_2.5.1         
 [5] fastmap_1.2.0     tidyselect_1.2.1  xfun_0.46         magrittr_2.0.3   
 [9] glue_1.7.0        tibble_3.2.1      knitr_1.48        pkgconfig_2.0.3  
[13] htmltools_0.5.8.1 rmarkdown_2.27    dplyr_1.1.4       generics_0.1.3   
[17] lifecycle_1.0.4   cli_3.6.3         fansi_1.0.6       vctrs_0.6.5      
[21] compiler_4.4.1    purrr_1.0.2       rstudioapi_0.16.0 tools_4.4.1      
[25] pillar_1.9.0      evaluate_0.24.0   yaml_2.3.10       rlang_1.1.4   

# Set up data directory

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/home/steven/Documents/workspace/thesis/ch2/gis_data')
```

# R library setup

```{r}
library(sf)
library(tidyverse)
library(rcartocolor)
library(sfnetworks)
library(qgisprocess)

```


# Pre-process shapefiles for analysis

```{r}

na_bbox <- read_sf('processed/qgis_bbox.gpkg')

###

zm_occ <- read_csv('downloads/zebra_mussel_occurrences/NAS-Specimen-Download.csv')

zm_occ <- zm_occ %>% select(`Scientific Name`, Longitude, Latitude, Accuracy, Status) %>% drop_na()

zm_occ_sf <- st_as_sf(zm_occ, coords = c('Longitude', 'Latitude'),
                   crs = st_crs(4326)) 

zm_occ_sf <- st_intersection(zm_occ_sf, na_bbox)

write_sf(zm_occ_sf, 'processed/dreissnid_occurrences.gpkg')

road_network <- read_sf('downloads/us_dot_road_network/NTAD_North_American_Roads_2691167086962748384.gpkg')

road_network <- road_network %>% filter(SURFACE == 'Unpaved' | SURFACE == 'Paved')

road_network <- st_intersection(road_network, na_bbox)

road_network <- road_network %>% filter(ID != 554663 & 
                                          ID != 554677 &
                                          ID != 554668 &
                                          ID != 554678 &
                                          ID != 554667 &
                                          ID != 554666)


write_sf(road_network, 'processed/usdot_road_network.gpkg')

boundaries <- st_read('downloads/political_boundaries/political_boundaries/boundaries_p_2021_v3.shp')

boundaries <- st_make_valid(boundaries)

boundaries <- boundaries %>% st_transform(crs = 4326)

coastline <- boundaries %>% st_union() %>% st_as_sf() %>% st_make_valid()

boundaries <- st_intersection(boundaries, na_bbox)

coastline <- st_as_sf(coastline) 

coastline <- st_intersection(coastline, na_bbox)

write_sf(boundaries, 'processed/cec_political_boundaries.gpkg')

write_sf(coastline, 'processed/cec_coastline.gpkg')

pop_places <- read_sf('downloads/populated_places/popPlaces_v2.shp')

pop_places <- pop_places %>% st_make_valid %>% st_transform(crs = 4326)

pop_places <- pop_places %>% filter(COUNTRY == 'USA' | COUNTRY == 'CAN' | COUNTRY == 'MEX')

pop_places <- st_intersection(pop_places, na_bbox)

write_sf(pop_places, 'processed/cec_populated_places.gpkg')

lakes <- read_sf('downloads/lakes_and_rivers/northamerica_lakes_cec_2023.shp')

lakes <- lakes %>% st_transform(crs = 4326) %>% st_make_valid()

lakes <- st_intersection(lakes, na_bbox)

write_sf(lakes, 'processed/cec_lakes.gpkg')

# Canada metropolitan areas

canada_metro <- read_sf('downloads/canada_census_metropolitan_areas/lcma000a16a_e.shp')

canada_metro <- canada_metro %>% st_transform(crs = 4326) %>% st_make_valid()

write_sf(canada_metro, 'processed/canada_metro_areas.gpkg')

# US metropolitan areas

us_metro <- read_sf('downloads/tiger_metropolitan_statistical_areas/tl_2019_us_cbsa.shp')

us_metro <- us_metro %>% st_transform(crs = 4326) %>% st_make_valid()

write_sf(us_metro, 'processed/us_metro_areas.gpkg')

# CCI Lakes

cci_lakes <- read_sf('/mnt/023ede0b-f66f-4769-939d-421b38967538/common/thesis/ch1/env_data/cci_lakes/cci_auxiliary_data/lakescci_v2.0.2_data-availability.shp')

cci_lakes <- cci_lakes %>% st_set_crs(4326) %>% st_make_valid()

write_sf(cci_lakes, 'processed/cci_lakes.gpkg')

```

