# Zebra mussel network modeling script

# Contact information

Contact: Steven Brownlee
Email: steven.fr.brownlee@gmail.com
Date last revised: June 24 2024

# Session info

R version 4.4.1 (2024-06-14)
Platform: x86_64-redhat-linux-gnu
Running under: Nobara Linux 40 (GNOME Edition)

Matrix products: default
BLAS/LAPACK: FlexiBLAS OPENBLAS-OPENMP;  LAPACK version 3.11.0

locale:
 [1] LC_CTYPE=en_CA.UTF-8      LC_NUMERIC=C              LC_TIME=en_CA.utf8       
 [4] LC_COLLATE=en_CA.UTF-8    LC_MONETARY=en_CA.utf8    LC_MESSAGES=en_CA.UTF-8  
 [7] LC_PAPER=en_CA.utf8       LC_NAME=C                 LC_ADDRESS=C             
[10] LC_TELEPHONE=C            LC_MEASUREMENT=en_CA.utf8 LC_IDENTIFICATION=C      

time zone: America/Vancouver
tzcode source: system (glibc)

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

loaded via a namespace (and not attached):
 [1] digest_0.6.36     tidyr_1.3.1       utf8_1.2.4        R6_2.5.1         
 [5] fastmap_1.2.0     tidyselect_1.2.1  xfun_0.46         magrittr_2.0.3   
 [9] glue_1.7.0        tibble_3.2.1      knitr_1.48        pkgconfig_2.0.3  
[13] htmltools_0.5.8.1 rmarkdown_2.27    dplyr_1.1.4       generics_0.1.3   
[17] lifecycle_1.0.4   cli_3.6.3         fansi_1.0.6       vctrs_0.6.5      
[21] compiler_4.4.1    purrr_1.0.2       rstudioapi_0.16.0 tools_4.4.1      
[25] pillar_1.9.0      evaluate_0.24.0   yaml_2.3.10       rlang_1.1.4   

# Set up data directory

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/home/steven/Documents/common/thesis/ch2/gis_data/')
```

# R library setup

```{r}
library(sf)
library(tidyverse)
library(rcartocolor)
library(sfnetworks)
library(qgisprocess)
library(osmdata)

```

# Import shapefiles

```{r}
zm_occurrences <- read_sf('data_directory/dreissenid_occurrences.gpkg')

us_major_cities_centroid <- read_sf('data_directory/cec_populated_places.gpkg') 

land <- read_sf('data_directory/cec_coastline.gpkg') 

lakes <- read_sf('data_directory/cec_lakes.gpkg') 

usdot_road_network <- read_sf('data_directory/usdot_road_network.gpkg')

qgis_bbox <- read_sf('data_directory/qgis_bbox.gpkg')

```


# Pre-processing

```{r}

zm_occurrences_proj <- st_transform(zm_occurrences, crs = 5070)

zm_10 <- st_buffer(zm_occurrences_proj, dist = 10000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '10 km')

zm_20 <- st_buffer(zm_occurrences_proj, dist = 20000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '20 km')

zm_30 <- st_buffer(zm_occurrences_proj, dist = 30000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '30 km')

zm_40 <- st_buffer(zm_occurrences_proj, dist = 40000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '40 km')

zm_50 <- st_buffer(zm_occurrences_proj, dist = 50000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '50 km')
 


 
zm_buffer_collected <- rbind(zm_50, zm_40, zm_30, zm_20, zm_10)

rm(zm_50, zm_40, zm_30, zm_20, zm_10)

zm_buffer_collected$buffer_dist <- factor(zm_buffer_collected$buffer_dist, levels = c('50 km', '40 km', '30 km', '20 km', '10 km'))

zm_buffer_collected <- st_transform(zm_buffer_collected, crs = 4326)

write_sf(zm_buffer_collected, 'data_directory/zm_buffer.gpkg')

```

# Load zm buffer

```{r}

zm_buffer_collected <- read_sf('data_directory/zm_buffer.gpkg')

```


```{r}

qgis_bbox_transformed <- qgis_bbox %>% st_transform(crs = 5070)

overlay_grid <- st_make_grid(qgis_bbox_transformed, cellsize = 2000, what = 'polygons',
                             square = FALSE) %>% st_as_sf() %>% st_transform(crs = 4326)

write_sf(overlay_grid, 'data_directory/2km_hexagon_grid.gpkg')

```

```{r}

overlay_grid <- read_sf('data_directory/2km_hexagon_grid.gpkg')

overlay_selection <- qgis_run_algorithm(
  "native:extractbylocation",
  INPUT = 'data_directory/2km_hexagon_grid.gpkg',
  PREDICATE = 6,
  INTERSECT = 'data_directory/cec_coastline.gpkg',
  OUTPUT = 'data_directory/2km_hexagon_selection.gpkg'
)

overlay_selection <- read_sf('data_directory/2km_hexagon_selection.gpkg')

write_sf(overlay_selection, 'data_directory/2km_hexagon_selection.gpkg', driver = 'GPKG')

```

# Read hexagon grid


```{r}

overlay_selection <- read_sf('data_directory/2km_hexagon_selection.gpkg') %>% 
  rowid_to_column('uid') 

overlay_centroids <- st_centroid(overlay_selection) %>% mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2])

overlay_centroids_joined <- st_join(overlay_centroids, overlay_selection)

write_sf(overlay_centroids_joined, 'data_directory/2km_centroids_joined.gpkg')

overlay_centroids <- overlay_centroids_joined %>% select(uid.y, lon, lat) %>% rename(uid = uid.y) %>% 
  st_drop_geometry()

write_rds(overlay_centroids, 'data_directory/2km_centroids_table.rds')

```

# Extract data from OpenStreetMap

```{r}

overlay_centroids <- readRDS('data_directory/2km_centroids_table.rds')


entry_table <- tibble(uid = numeric(), lanes = numeric(), maxspeed = numeric(),
                      type = factor(levels = c('motorway', 'trunk', 'primary', 
                                             'secondary', 'tertiary', 
                                             'unclassified', 'residential')))

osm_colnames <- c('osm_id', 'name', 'highway', 'lanes', 'maxspeed')

# Iterate through each of x number of centroid coordinate pairs, query 
# OpenStreetMap database and retrieve all road objects within 1000 m radius of
# centroid - approximately equivalent to half if hexagon polygon diameter.

for (x in 1:5549984) {
  # Extract lat/lon information and OSM query
  query_coords <- overlay_centroids %>% filter(uid == x)
  q_lon <- as.numeric(query_coords$lon)
  q_lat <- as.numeric(query_coords$lat)
  osm_query <- opq_around(q_lon, q_lat, 
                          radius = 1000, 
                          timeout = 50, 
                          key = 'highway') %>%
  osmdata_sf()
  
  osm_query_filtered <- osm_query$osm_lines
  
  print(paste('Query', x, 'retrieved! Assessing...'))
  
  # Only proceed to next step if all of needed column types are present, ------
  # if not write 'NA' in entry table 
  
  if (all((osm_colnames) %in% colnames(osm_query_filtered)) == TRUE) {
    
    # Filter for desired road types
     osm_query_filtered <- osm_query_filtered %>% 
       filter(highway == 'motorway' | highway == 'trunk' | 
                highway == 'primary' | highway == 'secondary' | 
                highway == 'tertiary' | highway == 'unclassified' |
                highway == 'residential')
     
    # Retrieve lane numbers ---------------------------------------------------
      osm_query_filtered$lanes <- as.numeric(osm_query_filtered$lanes) 
    
    # Pre-process speed data to remove 'mph', harmonize speeds in km/h --------
    osm_query_filtered <- osm_query_filtered %>% 
        mutate(mph = grepl('mph', osm_query_filtered$maxspeed))
    osm_query_filtered$maxspeed <- str_replace(osm_query_filtered$maxspeed, 
                                               'mph', '')
    osm_query_filtered$maxspeed <- as.numeric(osm_query_filtered$maxspeed)
    osm_query_filtered <- osm_query_filtered %>% 
        mutate(max_speed = case_when(mph == TRUE ~ maxspeed * 1.609344,
                                    mph == FALSE ~ maxspeed)) %>% 
        select(-maxspeed)
    osm_query_maxspeed <- osm_query_filtered %>% select(lanes) %>% 
        drop_na() %>% max()
    
    # Retrieve road types and set as factor based on capacity -----------------
    osm_query_filtered$type <- factor(osm_query_type$type, 
                                  levels = c('motorway', 'trunk', 'primary', 
                                             'secondary', 'tertiary', 
                                             'unclassified', 'residential'))
    osm_query_type <- osm_query_filtered %>% select(type) %>% drop_na()
    
    #--------------------------------------------------------------------------
    # Add entries to entry table
    entry_table <- entry_table %>% add_row(uid = x, 
                          lanes = osm_query_lanes, 
                          maxspeed = osm_query_maxspeed,
                          type = osm_query_type)
    print(paste('Completed:', x))
    rm(query_coords, q_lon, q_lat, osm_query, osm_query_filtered, osm_query_lanes,
      osm_query_maxspeed)
    
    #--------------------------------------------------------------------------
  } else {
    # Create empty entry if no useable data is present.
    entry_table <- entry_table %>% add_row(uid = x, 
                          lanes = NA, 
                          maxspeed = NA,
                          type = NA)
    print(paste('No useable data at:', x, 'proceeding!'))
  }
  }
```
# 

```{r}

#usdot_road_network <- usdot_road_network %>% filter(SURFACE == 'Unpaved' | SURFACE == 'Paved')

overlay_usdot <- st_join(overlay_selection, usdot_road_network, join = st_intersects)

overlay_usdot <- overlay_usdot %>% drop_na()

write_sf(overlay_usdot,'data_directory/2km_hexagon_joined.gpkg')

```

```{r}
overlay_usdot <- read_sf('data_directory/2km_hexagon_joined.gpkg')

```


```{r}

zm_bbox <- st_bbox(north_america_bbox)


buffer_plot <- ggplot() + geom_sf(data = land, fill = 'gray70', colour = 'gray15') +
  geom_sf(data = large_rivers, colour = 'gray60') +
  geom_sf(data = small_rivers, colour = 'gray60', alpha = 0.3) +
    geom_sf(data = large_lakes, fill = 'gray50', colour = 'gray15') +
  geom_sf(data = small_lakes, fill = 'gray50', colour = 'gray15') +
  geom_sf(data = zm_buffer_collected_test, aes(fill = buffer_dist)) +
  scale_fill_carto_d(palette = 'TealGrn', direction = -1, name = 'Buffer distance \n in km') +
  theme_dark(base_family = extrafont::choose_font('Cantarell Light'), 
               base_size = 12) +
      coord_sf(xlim = c(zm_bbox[[1]], zm_bbox[[3]]),
             ylim = c(zm_bbox[[2]], zm_bbox[[4]]), 
             expand = FALSE) 



```

```{r}

overlay_selection <- read_sf('data_directory/2km_hexagon_selection.gpkg')




```


Run:

usethis::edit_r_environ()

OSMEXT_DOWNLOAD_DIRECTORY=/home/steven/Documents/common/thesis/ch2/gis_data/downloads/open_street_map_r

```{r}


overlay_selection <- read_sf('data_directory/2km_hexagon_selection.gpkg')

test_filter <- overlay_selection %>% filter(uid == 496538)

            
              
      

canada <- oe_vectortranslate('downloads/open_street_map_r/geofabrik_canada-latest.osm.pbf')

canada_gpkg <- oe_read('downloads/open_street_map_r/geofabrik_canada-latest.gpkg')

canada_gpkg_test <- canada_gpkg %>% 
  filter(highway == "motorway" | highway == "tertiary" | highway == "residential" | 
           highway == "primary" | highway == "unclassified" | highway ==  "road" |
           highway == "trunk" | highway == "secondary" | highway == "motorway_link")

write_sf(canada_gpk_test, 'data_directory/')

bc <- oe_get('British Columbia', max_file_size = 4e+10, 
                 query = "SELECT * FROM 'lines' WHERE highway = 'motorway') 
                 OR highway = 'trunk' OR highway = 'primary' OR highway = 'secondary' OR highway = 'tertiary' OR highway = 'unclassified'  OR highway = 'residential' OR highway = 'motorway_link' OR highway = 'trunk_link' OR highway = 'primary_link' OR highway = 'secondary_link' OR highway = 'tertiary_link' OR highway = 'service' OR highway = 'track'")

```

```{r}
    if (length(row_number(osm_query_filtered)) > 0 ) {
      osm_query_filtered$lanes <- as.numeric(osm_query_filtered$lanes) 
      osm_query_filtered$type <- factor(osm_query_type$type, 
                                  levels = c('motorway', 'trunk', 'primary', 
                                             'secondary', 'tertiary', 
                                             'unclassified', 'residential'))
      osm_query_filtered <- osm_query_filtered %>% 
        mutate(mph = grepl('mph', osm_query_filtered$maxspeed))
      osm_query_filtered$maxspeed <- str_replace(osm_query_filtered$maxspeed, 'mph', '')
      osm_query_filtered$maxspeed <- as.numeric(osm_query_filtered$maxspeed)
      osm_query_filtered <- osm_query_filtered %>% 
        mutate(max_speed = case_when(mph == TRUE ~ maxspeed * 1.609344,
                                    mph == FALSE ~ maxspeed)) %>% 
        select(-maxspeed)
      osm_query_maxspeed <- osm_query_filtered %>% select(lanes) %>% 
        drop_na() %>% max()
      osm_query_type <- osm_query_filtered %>% select(type) %>% drop_na()
 
    } else {
      osm_query_lanes <- NA
      osm_query_maxspeed <- NA
      osm_query_type <- NA
    }
    entry_table %>% add_row(uid = x, 
                          lanes = osm_query_lanes, 
                          maxspeed = osm_query_maxspeed,
                          type = osm_query_type)
    print(paste('Completed:', x))
    rm(query_coords, q_lon, q_lat, osm_query, osm_query_filtered, osm_query_lanes,
      osm_query_maxspeed)
    }
    
  } else {
 entry_table %>% add_row(uid = x, 
                          lanes = NA, 
                          maxspeed = NA,
                          type = NA)
    
    
  }
```

