# Zebra mussel network modeling script

# Contact information

Contact: Steven Brownlee
Email: steven.fr.brownlee@gmail.com
Date last revised: June 24 2024

# Session info

R version 4.4.1 (2024-06-14)
Platform: x86_64-redhat-linux-gnu
Running under: Nobara Linux 40 (GNOME Edition)

Matrix products: default
BLAS/LAPACK: FlexiBLAS OPENBLAS-OPENMP;  LAPACK version 3.11.0

locale:
 [1] LC_CTYPE=en_CA.UTF-8      LC_NUMERIC=C              LC_TIME=en_CA.utf8       
 [4] LC_COLLATE=en_CA.UTF-8    LC_MONETARY=en_CA.utf8    LC_MESSAGES=en_CA.UTF-8  
 [7] LC_PAPER=en_CA.utf8       LC_NAME=C                 LC_ADDRESS=C             
[10] LC_TELEPHONE=C            LC_MEASUREMENT=en_CA.utf8 LC_IDENTIFICATION=C      

time zone: America/Vancouver
tzcode source: system (glibc)

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

loaded via a namespace (and not attached):
 [1] digest_0.6.36     tidyr_1.3.1       utf8_1.2.4        R6_2.5.1         
 [5] fastmap_1.2.0     tidyselect_1.2.1  xfun_0.46         magrittr_2.0.3   
 [9] glue_1.7.0        tibble_3.2.1      knitr_1.48        pkgconfig_2.0.3  
[13] htmltools_0.5.8.1 rmarkdown_2.27    dplyr_1.1.4       generics_0.1.3   
[17] lifecycle_1.0.4   cli_3.6.3         fansi_1.0.6       vctrs_0.6.5      
[21] compiler_4.4.1    purrr_1.0.2       rstudioapi_0.16.0 tools_4.4.1      
[25] pillar_1.9.0      evaluate_0.24.0   yaml_2.3.10       rlang_1.1.4   

# Set up data directory

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/home/steven/Documents/common/thesis/ch2/gis_data/')
```

# R library setup

```{r}
library(sf)
library(tidyverse)
library(rcartocolor)
library(sfnetworks)
library(qgisprocess)
library(future)
library(doFuture)
library(purrr)
library(furrr)

```

# Import shapefiles

```{r}
zm_occurrences <- read_sf('data_directory/dreissenid_occurrences.gpkg')

us_major_cities_centroid <- read_sf('data_directory/cec_populated_places.gpkg') 

land <- read_sf('data_directory/cec_coastline.gpkg') 

lakes <- read_sf('data_directory/cec_lakes.gpkg') 

usdot_road_network <- read_sf('data_directory/usdot_road_network.gpkg')

qgis_bbox <- read_sf('data_directory/qgis_bbox.gpkg')

```


# Pre-processing

```{r}

zm_occurrences_proj <- st_transform(zm_occurrences, crs = 5070)

zm_10 <- st_buffer(zm_occurrences_proj, dist = 10000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '10 km')

zm_20 <- st_buffer(zm_occurrences_proj, dist = 20000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '20 km')

zm_30 <- st_buffer(zm_occurrences_proj, dist = 30000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '30 km')

zm_40 <- st_buffer(zm_occurrences_proj, dist = 40000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '40 km')

zm_50 <- st_buffer(zm_occurrences_proj, dist = 50000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '50 km')
 


 
zm_buffer_collected <- rbind(zm_50, zm_40, zm_30, zm_20, zm_10)

rm(zm_50, zm_40, zm_30, zm_20, zm_10)

zm_buffer_collected$buffer_dist <- factor(zm_buffer_collected$buffer_dist, levels = c('50 km', '40 km', '30 km', '20 km', '10 km'))

zm_buffer_collected <- st_transform(zm_buffer_collected, crs = 4326)

write_sf(zm_buffer_collected, 'data_directory/zm_buffer.gpkg')

```

# Load zm buffer

```{r}

zm_buffer_collected <- read_sf('data_directory/zm_buffer.gpkg')

```


```{r}

qgis_bbox_transformed <- qgis_bbox %>% st_transform(crs = 5070)

overlay_grid <- st_make_grid(qgis_bbox_transformed, cellsize = 2000, what = 'polygons',
                             square = FALSE) %>% st_as_sf() %>% st_transform(crs = 4326)

write_sf(overlay_grid, 'data_directory/2km_hexagon_grid.gpkg')

```

```{r}

overlay_grid <- read_sf('data_directory/2km_hexagon_grid.gpkg')

overlay_selection <- qgis_run_algorithm(
  "native:extractbylocation",
  INPUT = 'data_directory/2km_hexagon_grid.gpkg',
  PREDICATE = 6,
  INTERSECT = 'data_directory/cec_coastline.gpkg',
  OUTPUT = 'data_directory/2km_hexagon_selection.gpkg'
)

overlay_selection <- read_sf('data_directory/2km_hexagon_selection.gpkg')

write_sf(overlay_selection, 'data_directory/2km_hexagon_selection.gpkg', driver = 'GPKG')

```

# Read hexagon grid


```{r}

overlay_selection <- read_sf('data_directory/2km_hexagon_selection.gpkg') %>% 
  rowid_to_column('uid') 

overlay_centroids <- st_centroid(overlay_selection) %>% mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2])

overlay_centroids_joined <- st_join(overlay_centroids, overlay_selection)

write_sf(overlay_centroids_joined, 'data_directory/2km_centroids_joined.gpkg')

overlay_centroids <- overlay_centroids_joined %>% select(uid.y, lon, lat) %>% rename(uid = uid.y) %>% 
  st_drop_geometry()

write_rds(overlay_centroids, 'data_directory/2km_centroids_table.rds')

```

# Extract data from OpenStreetMap

```{r}

overlay_centroids <- readRDS('data_directory/2km_centroids_table.rds')

```
# d

# New workflow

```{r}

na_data <- read_sf('data_directory/na_highway_filtered_final.gpkg')

na_data <- na_data %>% select(-waterway, -aerialway, -barrier, -man_made, -z_order) 

na_data <- st_centroid(na_data)

na_data <- st_join(na_data, hexagon_dissolved)

write_sf(na_data, 'data_directory/na_highway_grouped.gpkg')

```



```{r}

hexagon_dissolved <- read_sf('data_directory/2km_hexagon_dissolved.gpkg')

na_data <- read_sf('data_directory/na_highway_filtered_final.gpkg') %>% 
  select(-waterway, -aerialway, -barrier, -man_made, -z_order) %>%
  group_by(name) %>% 
  mutate(group_number = cur_group_id())


###

divided_dataset <- tibble(x = seq(1:2487683))

num_groups = 500

divided_dataset <- tibble(x = seq(1:500))

num_groups = 3

divided_dataset <- divided_dataset %>% mutate(grouping = (row_number()-1) %/% (n()/num_groups)) %>% 
   group_by(grouping) %>% 
  summarize(min_x = min(x), max_x = max(x)) %>% select(-grouping)


sequence_start = pull(divided_dataset, min_x)

sequence_end = pull(divided_dataset, max_x)

for (i in 1:500)  {
  group_slice <- na_data %>% 
    filter(group_number >= sequence_start[i] & group_number <= sequence_end[i]) %>% 
    st_transform(crs = 'ESRI:102008')
  write_sf(group_slice, paste0('scratch/group_slice_', i, '.gpkg'))
  print(paste('Completed:', i))
}

osm_slice_list = list.files('/home/steven/Documents/common/thesis/ch2/gis_data/', 
                      pattern = 'group_slice_*',
                      full.names = TRUE,
                      recursive = TRUE)

osm_table <- tibble(grouping_id = numeric(),
                    osm_id = character(),
                    name = character(),
                    highway = character(),
                    other_tags = character(),
                    group_number = numeric(),
                    lon = numeric(),
                    lat = numeric())
 
osm_processor <- function(linestring) {
  osm_linestring <- read_sf(linestring)
  group_indices <- unique(osm_linestring$group_number)
  osm_list = vector('list', length = length(group_indices))
  counter = 1

    for (i in group_indices) {
      suppressWarnings({
        osm_segmented <-  qgis_run_algorithm("native:splitlinesbylength",
                     INPUT = filter(osm_linestring, 
                                    group_number == i),
                     LENGTH = 1000) %>% st_as_sf()
        osm_points <- qgis_run_algorithm("native:pointsalonglines",
                                 INPUT = osm_segmented,
                                 DISTANCE = 1000) %>% 
          st_as_sf() %>% 
         mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2]) %>% 
        st_drop_geometry() %>% 
        select(group_number, osm_id, name, highway, 
               other_tags, lon, lat)
        osm_list[[counter]] <- osm_points
        print(paste('Completed:', counter))
        counter = counter + 1
        qgis_clean_tmp()
        gc()
      })
  }
  osm_dataframe <- bind_rows(osm_list)
  print(paste('Completed slice!'))
  return(osm_dataframe)
}


plan(multisession, workers = 6)

slice_one <- osm_slice_list[1:167]
slice_two <- osm_slice_list[168:334]
slice_three <- osm_slice_list[334:500]

osm_large_list <- future_map(osm_processor(osm_slice_list), .progress = TRUE)

###

osm_data_list = list.files('scratch/', 
                      pattern = 'osm_points_500m_intervals_*',
                      full.names = TRUE,
                      recursive = TRUE)

for (i in 1:length(osm_data_list))  {
 osm_point_intermediate <- read_sf(osm_data_list[i]) %>% 
   st_transform(crs = 4326)
 osm_point_intermediate <- st_join(osm_point_intermediate, hexagon_dissolved) 
 write_sf(osm_point_intermediate, osm_data_list[i])
 print(paste('Written:', i))
}


```


```{r}

combined_osm_table <- tibble(osm_id = numeric(),
                             name = character(),
                             highway = character(),
                             other_tags = character(),
                             group_id = numeric(),
                             lon = numeric(),
                             lat = numeric())


osm_data_list = list.files('scratch/', 
                      pattern = 'osm_points_500m_intervals_*',
                      full.names = TRUE,
                      recursive = TRUE)


for (i in 1:length(osm_data_list)) {
  
  osm_subset <- read_sf(osm_data_list[i])
  osm_subset <- osm_subset %>% mutate(lon = sf::st_coordinates(.)[,1],
                lat = sf::st_coordinates(.)[,2]) %>% 
    st_drop_geometry()
  combined_osm_table <- rbind(combined_osm_table, osm_subset)
  print(paste('Completed:', i))
}

#write_rds(combined_osm_table, 'data_directory/combined_osm_table.rds')

#combined_osm_table <- readRDS('data_directory/combined_osm_table.rds') 

combined_osm_table <- combined_osm_table %>% filter(!group_id %in% c(NA)) %>% 
  select(-group_split) %>% rowid_to_column('point_id')

osm_table_ids <- combined_osm_table %>% select(point_id, osm_id, group_id, lon, name, lat)

osm_table_lanes <- combined_osm_table %>% filter(grepl('lanes', other_tags)) %>% 
    separate_longer_delim(other_tags, delim = ',') %>%  
    filter(grepl('"lanes"=>', other_tags)) %>%  
    mutate(across('other_tags', str_replace, '"lanes"=>', '')) %>% 
    mutate(across('other_tags', str_replace, '"', '')) %>% 
    mutate(across('other_tags', str_replace, '"', '')) %>% 
    rename(lanes = other_tags) %>% select(point_id, osm_id, name, lanes) %>% 
    st_drop_geometry()

osm_table_lanes <- osm_table_lanes

osm_table_lanes$lanes <- sub("\\;.*", "", osm_table_lanes$lanes)

osm_table_lanes$lanes <- gsub("[^0-9]", NA, osm_table_lanes$lanes)

osm_table_lanes <- osm_table_lanes %>% drop_na()

osm_table_lanes$lanes <- as.numeric(osm_table_lanes$lanes)

####

osm_table_maxspeed <- combined_osm_table %>% filter(grepl('maxspeed', other_tags)) %>% 
    separate_longer_delim(other_tags, delim = ',')  %>% 
    filter(grepl('"maxspeed"=>', other_tags)) %>% 
    mutate(across('other_tags', str_replace, '"maxspeed"=>', ''))%>% 
    rename(max_speed = other_tags)  %>% 
    filter(!grepl('unposted', max_speed)) %>% 
    filter(!grepl('slow', max_speed)) %>% 
    filter(!grepl('implicit', max_speed)) %>% 
    filter(!grepl('default', max_speed)) %>% 
     filter(!grepl('signals', max_speed)) %>% 
     filter(!grepl('Variable', max_speed)) %>% 
    select(point_id, osm_id, max_speed)
  
osm_table_maxspeed_kmh <- osm_table_maxspeed %>% 
  filter(!grepl('mph', max_speed) & !grepl('MPH', max_speed)) %>% 
    mutate(across('max_speed', str_replace, '"', '')) %>% 
    mutate(across('max_speed', str_replace, '"', ''))

osm_table_maxspeed_kmh$max_speed <- sub("\\;.*", "", osm_table_maxspeed_kmh$max_speed)

osm_table_maxspeed_kmh$max_speed <- gsub("[^0-9.-]", NA, osm_table_maxspeed_kmh$max_speed) 

osm_table_maxspeed_kmh$max_speed <- as.numeric(osm_table_maxspeed_kmh$max_speed)
  
osm_table_maxspeed_kmh <- osm_table_maxspeed_kmh %>% drop_na()

###

osm_table_maxspeed_mph <- osm_table_maxspeed %>% filter(grepl('mph', max_speed)) %>% 
  mutate(across('max_speed', str_replace, ' mph', '')) %>%
  mutate(across('max_speed', str_replace, 'mph', ''))

osm_table_maxspeed_mph$max_speed <- sub("\\;.*", "", osm_table_maxspeed_mph$max_speed)

osm_table_maxspeed_mph <- osm_table_maxspeed_mph %>% 
  mutate(across('max_speed', str_replace, '"', '')) %>% 
  mutate(across('max_speed', str_replace, '"', '')) 

osm_table_maxspeed_mph$max_speed <- as.numeric(osm_table_maxspeed_mph$max_speed)

osm_table_maxspeed_mph <- osm_table_maxspeed_mph %>% drop_na()

osm_table_maxspeed_mph <- osm_table_maxspeed_mph %>% 
    mutate(max_speed_kmh = max_speed*1.609344) %>% 
    select(point_id, osm_id, max_speed_kmh) %>% 
    rename(max_speed = max_speed_kmh)

osm_table_maxspeed_combined <- rbind(osm_table_maxspeed_mph, osm_table_maxspeed_kmh)

osm_table_maxspeed_combined <- osm_table_maxspeed_combined %>% 
  filter(max_speed < 120)


###

osm_table_type <- combined_osm_table %>% select(point_id, osm_id, highway) %>% rename(type = highway)

osm_table_type$type <- factor(osm_table_type$type, 
                                  levels = c('residential', 'unclassified', 
                                               'tertiary', 'secondary', 
                                               'primary', 'trunk', 'motorway'))

osm_filtered_data <- left_join(osm_table_ids, osm_table_type, 
                               by = 'point_id') %>% drop_na()

osm_filtered_data <- left_join(osm_filtered_data, osm_table_lanes, 
                               by = 'point_id')

osm_filtered_data <- left_join(osm_filtered_data, osm_table_maxspeed_combined, 
                               by = 'point_id')

#osm_filtered_data <- osm_filtered_data %>% drop_na()

osm_filtered_data <- st_as_sf(osm_filtered_data, coords = c('lon', 'lat'), 
                              crs = st_crs(4326))


write_sf(osm_filtered_data, 'data_directory/osm_filtered_points.gpkg')


```


```{r}

hexagon_selection <- read_sf('data_directory/2km_hexagon_selection.gpkg') %>% rowid_to_column()

na_points_total <- read_sf('data_directory/osm_points_final.gpkg') 

divided_dataset <- tibble(x = seq(1:5549984))

num_groups = 40

divided_dataset <- divided_dataset %>% mutate(grouping = (row_number()-1) %/% (n()/num_groups)) %>% 
   group_by(grouping) %>% 
  summarize(min_x = min(x), max_x = max(x)) %>% select(-grouping)


sequence_start = pull(divided_dataset, min_x)

sequence_end = pull(divided_dataset, max_x)

hexagon_selection <- hexagon_selection %>% mutate(group_id = case_when(
  rowid >= sequence_start[1] & rowid <= sequence_end[1] ~ 1,
  rowid >= sequence_start[2] & rowid <= sequence_end[2] ~ 2,
  rowid >= sequence_start[3] & rowid <= sequence_end[3] ~ 3,
  rowid >= sequence_start[4] & rowid <= sequence_end[4] ~ 4,
  rowid >= sequence_start[5] & rowid <= sequence_end[5] ~ 5,
  rowid >= sequence_start[6] & rowid <= sequence_end[6] ~ 6,
  rowid >= sequence_start[7] & rowid <= sequence_end[7] ~ 7,
  rowid >= sequence_start[8] & rowid <= sequence_end[8] ~ 8,
  rowid >= sequence_start[9] & rowid <= sequence_end[9] ~ 9,
  rowid >= sequence_start[10] & rowid <= sequence_end[10] ~ 10,
  rowid >= sequence_start[11] & rowid <= sequence_end[11] ~ 11,
  rowid >= sequence_start[12] & rowid <= sequence_end[12] ~ 12,
  rowid >= sequence_start[13] & rowid <= sequence_end[13] ~ 13,
  rowid >= sequence_start[14] & rowid <= sequence_end[14] ~ 14,
  rowid >= sequence_start[15] & rowid <= sequence_end[15] ~ 15,
  rowid >= sequence_start[16] & rowid <= sequence_end[16] ~ 16,
  rowid >= sequence_start[17] & rowid <= sequence_end[17] ~ 17,
  rowid >= sequence_start[18] & rowid <= sequence_end[18] ~ 18,
  rowid >= sequence_start[19] & rowid <= sequence_end[19] ~ 19,
  rowid >= sequence_start[20] & rowid <= sequence_end[20] ~ 20,
  rowid >= sequence_start[21] & rowid <= sequence_end[21] ~ 21,
  rowid >= sequence_start[22] & rowid <= sequence_end[22] ~ 22,
  rowid >= sequence_start[23] & rowid <= sequence_end[23] ~ 23,
  rowid >= sequence_start[24] & rowid <= sequence_end[24] ~ 24,
  rowid >= sequence_start[25] & rowid <= sequence_end[25] ~ 25,
  rowid >= sequence_start[26] & rowid <= sequence_end[26] ~ 26,
  rowid >= sequence_start[27] & rowid <= sequence_end[27] ~ 27,
  rowid >= sequence_start[28] & rowid <= sequence_end[28] ~ 28,
  rowid >= sequence_start[29] & rowid <= sequence_end[29] ~ 29,
  rowid >= sequence_start[30] & rowid <= sequence_end[30] ~ 30,
  rowid >= sequence_start[31] & rowid <= sequence_end[31] ~ 31,
  rowid >= sequence_start[32] & rowid <= sequence_end[32] ~ 32,
  rowid >= sequence_start[33] & rowid <= sequence_end[33] ~ 33,
  rowid >= sequence_start[34] & rowid <= sequence_end[34] ~ 34,
  rowid >= sequence_start[35] & rowid <= sequence_end[35] ~ 35,
  rowid >= sequence_start[36] & rowid <= sequence_end[36] ~ 36,
  rowid >= sequence_start[37] & rowid <= sequence_end[37] ~ 37,
  rowid >= sequence_start[38] & rowid <= sequence_end[38] ~ 38,
  rowid >= sequence_start[39] & rowid <= sequence_end[39] ~ 39,
  rowid >= sequence_start[40] & rowid <= sequence_end[40] ~ 40))

write_sf(hexagon_selection, 'data_directory/2km_hexagon_selection_grouped.gpkg') 

hexagon_simplified <- hexagon_selection %>% group_by(group_id) %>% summarize()

na_points_grouped <- st_join(na_points_total, hexagon_simplified, join = st_intersects,
                             left = TRUE)

write_sf(na_points_grouped, 'data_directory/osm_points_grouped.gpkg')

###

osm_grid <- read_sf('data_directory/2km_hexagon_selection_grouped.gpkg')
osm_points <- read_sf('data_directory/osm_points_grouped.gpkg')

# Note: fix osm_points converting max_speed to character

osm_points$max_speed <- as.numeric(osm_points$max_speed)

###

combined_table <- tibble(rowid = numeric(),
                         max_speed = numeric(),
                         type = character(),
                         lanes = numeric())


for (i in seq(1:40)) {
  hexagon_filtered <- osm_grid %>% filter(group_id == i)
  na_points_filtered <- osm_points %>% filter(group_id == i)
  joined_points <- st_join(na_points_filtered, hexagon_filtered, 
                               left = TRUE,
                               join = st_intersects) %>% 
        st_drop_geometry()
      
  hexagon_sum <- joined_points %>% group_by(rowid) %>% 
        summarize_at(c('max_speed', 'type', 'lanes'),
                     max, na.rm = TRUE)
  print(nrow(hexagon_sum))
 # combined_table <- rbind(combined_table, hexagon_sum)
  print(paste('Completed:', i))
}

write_csv(combined_table, 'data_directory/osm_grid_centroids.csv')

hexagon_osm_final <- inner_join(hexagon_selection, combined_table, by = 'rowid')

write_sf(hexagon_osm_final, 'data_directory/osm_hexagons.test.gpkg')

###

hexagon_filtered <- osm_grid %>% filter(group_id == a)
  na_points_filtered <- osm_points %>% filter(group_id == a)
  joined_points <- st_join(na_points_filtered, hexagon_filtered, 
                               left = TRUE,
                               join = st_intersects) %>% 
        st_drop_geometry()
      
  hexagon_sum <- joined_points %>% group_by(rowid) %>% 
        summarize_at(c('max_speed', 'type', 'lanes'),
                     max, na.rm = TRUE)
  print(nrow(hexagon_sum))
  combined_table <- rbind(combined_table, hexagon_sum)
  print(nrow(hexagon_sum))
  print(paste('Completed:', i))






plan(multicore, workers = 8)
options(future.globals.maxSize = 2e+9)
registerDoFuture()


combined_table <- foreach(a = 1:40,
                          .combine = rbind) %dofuture% {
                            
      hexagon_filtered <- osm_grid %>% filter(group_id == a)
      na_points_filtered <- osm_points %>% filter(group_id == a)
      joined_points <- st_join(na_points_filtered, hexagon_filtered, 
                               left = TRUE,
                               join = st_intersects) %>% 
        st_drop_geometry()
      
      hexagon_sum <- joined_points %>% group_by(rowid) %>% 
        summarize_at(c('max_speed', 'type', 'lanes'),
                     max, na.rm = TRUE)
      return(hexagon_sum)
      }


```



```{r}

zm_bbox <- st_bbox(north_america_bbox)


buffer_plot <- ggplot() + geom_sf(data = land, fill = 'gray70', colour = 'gray15') +
  geom_sf(data = large_rivers, colour = 'gray60') +
  geom_sf(data = small_rivers, colour = 'gray60', alpha = 0.3) +
    geom_sf(data = large_lakes, fill = 'gray50', colour = 'gray15') +
  geom_sf(data = small_lakes, fill = 'gray50', colour = 'gray15') +
  geom_sf(data = zm_buffer_collected_test, aes(fill = buffer_dist)) +
  scale_fill_carto_d(palette = 'TealGrn', direction = -1, name = 'Buffer distance \n in km') +
  theme_dark(base_family = extrafont::choose_font('Cantarell Light'), 
               base_size = 12) +
      coord_sf(xlim = c(zm_bbox[[1]], zm_bbox[[3]]),
             ylim = c(zm_bbox[[2]], zm_bbox[[4]]), 
             expand = FALSE) 



```
