# Zebra mussel network modeling script

# Contact information

Contact: Steven Brownlee
Email: steven.fr.brownlee@gmail.com
Date last revised: June 24 2024

# Session info

R version 4.4.1 (2024-06-14)
Platform: x86_64-redhat-linux-gnu
Running under: Nobara Linux 40 (GNOME Edition)

Matrix products: default
BLAS/LAPACK: FlexiBLAS OPENBLAS-OPENMP;  LAPACK version 3.11.0

locale:
 [1] LC_CTYPE=en_CA.UTF-8      LC_NUMERIC=C              LC_TIME=en_CA.utf8       
 [4] LC_COLLATE=en_CA.UTF-8    LC_MONETARY=en_CA.utf8    LC_MESSAGES=en_CA.UTF-8  
 [7] LC_PAPER=en_CA.utf8       LC_NAME=C                 LC_ADDRESS=C             
[10] LC_TELEPHONE=C            LC_MEASUREMENT=en_CA.utf8 LC_IDENTIFICATION=C      

time zone: America/Vancouver
tzcode source: system (glibc)

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

loaded via a namespace (and not attached):
 [1] digest_0.6.36     tidyr_1.3.1       utf8_1.2.4        R6_2.5.1         
 [5] fastmap_1.2.0     tidyselect_1.2.1  xfun_0.46         magrittr_2.0.3   
 [9] glue_1.7.0        tibble_3.2.1      knitr_1.48        pkgconfig_2.0.3  
[13] htmltools_0.5.8.1 rmarkdown_2.27    dplyr_1.1.4       generics_0.1.3   
[17] lifecycle_1.0.4   cli_3.6.3         fansi_1.0.6       vctrs_0.6.5      
[21] compiler_4.4.1    purrr_1.0.2       rstudioapi_0.16.0 tools_4.4.1      
[25] pillar_1.9.0      evaluate_0.24.0   yaml_2.3.10       rlang_1.1.4   

# Set up data directory

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/home/steven/Documents/workspace/thesis/ch2/gis_data/')
```

# R library setup

```{r}
library(sf)
library(tidyverse)
library(rcartocolor)
library(sfnetworks)
library(qgisprocess)
library(future)
library(future.callr)
library(purrr)
library(furrr)
library(progressr)

```

# Import shapefiles

```{r}
zm_occurrences <- read_sf('processed/dreissenid_occurrences.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

us_metro <- read_sf('processed/us_metro_areas.gpkg')%>% 
  st_transform(crs = 'ESRI:102008')

canada_metro <- read_sf('processed/canada_metro_areas.gpkg')%>% 
  st_transform(crs = 'ESRI:102008')

land <- read_sf('processed/shared_north_america_coastline.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

lakes <- read_sf('processed/cec_lakes.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

usdot_road_network <- read_sf('processed/usdot_road_network.gpkg')%>% 
  st_transform(crs = 'ESRI:102008')

qgis_bbox <- read_sf('processed/qgis_bbox.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

```


# Pre-processing

```{r}
# Create buffers around ZM occurrences

zm_10 <- st_buffer(zm_occurrences, dist = 10000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '10 km')

zm_20 <- st_buffer(zm_occurrences, dist = 20000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '20 km')

zm_30 <- st_buffer(zm_occurrences, dist = 30000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '30 km')

zm_40 <- st_buffer(zm_occurrences, dist = 40000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '40 km')

zm_50 <- st_buffer(zm_occurrences, dist = 50000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '50 km')

zm_buffer_collected <- rbind(zm_50, zm_40, zm_30, zm_20, zm_10)

rm(zm_50, zm_40, zm_30, zm_20, zm_10)

zm_buffer_collected$buffer_dist <- factor(zm_buffer_collected$buffer_dist, levels = c('50 km', '40 km', '30 km', '20 km', '10 km'))

zm_buffer_collected <- st_transform(zm_buffer_collected, crs = 4326)

write_sf(zm_buffer_collected, 'processed/zm_buffer.gpkg')

# Merge US and Canada metro areas

us_metro <- us_metro %>% select(NAME) %>% rename(metro_area = NAME)

canada_metro <- canada_metro %>% select(CMANAME) %>% rename(metro_area = CMANAME)

north_america_metro <- rbind(us_metro, canada_metro)

# Create node network for USDOT roads
```

# Split road into manageable chunks

```{r}

usdot_road_network <- usdot_road_network %>% rownames_to_column('seg_id')


plan(callr, workers = 6)
options(future.globals.maxSize= 891289600)

node_generator <- function(sublist) {

for (x in sublist) {
  usdot_template <- tibble(x_coord = numeric(),
                         y_coord = numeric(),
                         country = character(),
                         usdot_id = numeric(),
                         jurisdiction = character(),
                         road_number = character(),
                         road_name = character(),
                         surface_type = character(),
                         lanes = numeric(),
                         speed_limit = numeric(),
                         class = numeric())
  
segment_filtered <- usdot_road_network %>% filter(seg_id == x)
length_extract = segment_filtered$LENGTH

if (length_extract <= 1) {
  sample_size = 5
} else if (1 < length_extract & length_extract <= 5 ) {
  sample_size = 20
} else if (5 < length_extract & length_extract <= 20) {
  sample_size = 100 
} else if (20 < length_extract & length_extract <= 50) {
  sample_size = 400 
} else if (50 < length_extract & length_extract <= 100) {
  sample_size = 700
} else if (100 < length_extract & length_extract <= 400) {
  sample_size = 2500
} else {sample_size = 2000} 

usdot_nodes <- st_sample(segment_filtered, size = sample_size, type = 'regular') %>% 
  st_cast("POINT") %>% 
  st_as_sf() %>% 
  mutate(x_coord = sf::st_coordinates(.)[,1],
         y_coord = sf::st_coordinates(.)[,2],
         usdot_id = segment_filtered$ID,
         country = segment_filtered$COUNTRY,
         jurisdiction = segment_filtered$JURISNAME,
         road_number = segment_filtered$ROADNUM,
         road_name = segment_filtered$ROADNAME,
         surface_type = segment_filtered$SURFACE,
         lanes = segment_filtered$LANES,
         speed_limit = segment_filtered$SPEEDLIM,
         class = segment_filtered$CLASS) %>% 
  st_drop_geometry()

usdot_nodes$x_coord <- as.numeric(usdot_nodes$x_coord)
usdot_nodes$y_coord <- as.numeric(usdot_nodes$y_coord)

usdot_template <- rbind(usdot_template, usdot_nodes) 

}
usdot_filled <- usdot_template
return(usdot_filled)
}

usdot_nodes <- future_map(1:716638, ~node_generator(.x))

###

usdot_nodes_2 <- bind_rows(usdot_nodes)

write_rds(usdot_nodes_2, 'outputs/node_folder/usdot_combined_2.rds')

rm(usdot_nodes_2, usdot_nodes)

usdot_combined <- read_rds('outputs/node_folder/usdot_combined_2.rds') 

usdot_combined <- usdot_combined %>% rowid_to_column('node_id')


inside_list = c(1, 500000,
                1000000,
                2000000,
                3000000,
                4000000,
                5000000,
                6000000,
                7000000,
                8000000,
                9000000,
                10000000,
                11000000)

outside_list= c(500000,
                1000000,
                2000000,
                3000000,
                4000000,
                5000000,
                6000000,
                7000000,
                8000000,
                9000000,
                10000000,
                11000000,
                11544435)

for (i in 1:13) {
  inside_index <- inside_list[i]
  outside_index <- outside_list[i]
  usdot_subset <- usdot_combined %>% 
    filter(node_id >= inside_index & node_id < outside_index)
   usdot_subset <- usdot_subset %>% 
     drop_na() %>% 
  st_as_sf(coords = c('x_coord', 'y_coord'), crs = st_crs('ESRI:102008'))
  write_sf(usdot_subset, paste0('outputs/node_folder/node_pieces/node_subset_', i, '.gpkg'))
  print(paste('Completed:', i))
}

###
```


# Join road data to grid
```{r}
#overlay_grid <- st_make_grid(qgis_bbox, cellsize = 2000,
#                             what = 'polygons', square = FALSE) %>% st_as_sf() %>% 
#  st_transform(crs = 'ESRI:102008')

#write_sf(overlay_grid, 'processed/2km_hexagon_grid.gpkg')

overlay_grid <- read_sf('processed/2km_hexagon_grid.gpkg')

overlay_grid <- overlay_grid %>% rowid_to_column('hex_id')

overlay_grid$hex_group <- as.numeric(cut_number(overlay_grid$hex_id, 1000))

for (x in 1:1000) {
  overlay_subset <- overlay_grid %>% filter(hex_group == x)
  write_sf(overlay_subset, paste0('outputs/overlay_grid_segments/hex_group_', x, '.gpkg'))
  print(paste('Completed:', x))
}


hex_list <- list.files('outputs/overlay_grid_segments/', 
                            pattern = '*.gpkg',
                            full.names = TRUE,
                            recursive = TRUE)

node_list <- list.files('outputs/node_folder/node_pieces/', 
                            pattern = '*.gpkg',
                            full.names = TRUE,
                            recursive = TRUE)


table_template <- tibble(hexid = numeric(),
                         node_id = numeric(),
                         usdot_id = numeric(),
                         country = numeric(),
                         jurisdiction = character(),
                         road_number = character(),
                         road_name = character(),
                         surface_type = character(),
                         lanes = numeric())

joiner <- function(sequence) {
for (x in sequence) {
  hex_segment  <- read_sf(hex_list[x]) %>% select(-hex_group)
  for (i in 1:13) {
    interior_nodes  <- read_sf(node_list[i])
    hex_interior <- st_join(hex_segment, interior_nodes) %>% st_drop_geometry()
    table_template <- rbind(table_template, hex_interior)
    print(paste('Interior done:', i))
}
  table_export <- table_template
  write_rds(table_export, paste0('outputs/overlay_grid_joined/overlay_grid_joined_', x, '.rds'))
  table_template <- tibble(hexid = numeric(),
                         node_id = numeric(),
                         usdot_id = numeric(),
                         country = numeric(),
                         jurisdiction = character(),
                         road_number = character(),
                         road_name = character(),
                         surface_type = character(),
                         lanes = numeric())
  return(table_export)
}
}

plan(callr, workers = 6)
options(future.globals.maxSize= 891289600)


joined <- future_map(1:1000, ~joiner(.x))

joined <- bind_rows(joined)

write_rds(joined, 'outputs/overlay_grid_joined/overlay_tables/joined_table.rds')

joined_table <- drop_na(joined)

write_rds(joined_table, 'outputs/overlay_grid_joined/overlay_tables/joined_table_na_dropped.rds')

joined_table <- read_rds('outputs/overlay_grid_joined/overlay_tables/joined_table_na_dropped.rds')

joined_table_summarized <- joined_table %>% select(hex_id, country, jurisdiction,
                                                   surface_type, surface_type,
                                                   lanes, speed_limit, class)

joined_table_summarized$surface_type <- case_match(joined_table_summarized$surface_type, 
                                            'Unpaved' ~ '1', 
                                            'Paved' ~ '2')

joined_table_summarized$surface_type <- as.numeric(joined_table_summarized$surface_type)

joined_table_summarized <- joined_table %>% 
  group_by(hex_id) %>% 
  summarize(surface_type = max(surface_type),
            lanes = max(lanes),
            speed_limit = max(speed_limit))

hex_final <- left_join(overlay_grid, joined_table_summarized, by = 'hex_id') %>% 
  drop_na()


write_sf(hex_final, 'outputs/network_components/hex_grid_final.gpkg')

```

# Edge generation

```{r}
hexagon_grid <- read_sf('outputs/network_components/hex_grid_final.gpkg') %>% 
  rowid_to_column('hex_rowid')

hex_centroids <- hexagon_grid %>% st_centroid() %>%  
  mutate(x = sf::st_coordinates(.)[,1],
         y = sf::st_coordinates(.)[,2]) %>% 
  select(hex_rowid, surface_type, lanes, speed_limit, x, y, geom)

#write_sf(hex_centroids, 'outputs/network_components/centroids.gpkg')

hex_centroids_aspatial <- hex_centroids %>% st_drop_geometry()

###

row_indices <- split(hex_centroids$hex_rowid, cut(hex_centroids$hex_rowid, breaks = 160))


###

edge_generator <- function(row_subset) {
  
  distance_array_template <- tibble(hex_rowid_1 = numeric(),
                                    x_1 = numeric(),
                                    y_1 = numeric(),
                                    hex_rowid_2 = numeric(),
                                    x_2 = numeric(),
                                    y_2 = numeric(),
                                    comb_hex_rowid = character())
  
  for (x in row_subset) {
    xval = eval(x)
    coord_1 <- hex_centroids %>% filter(hex_rowid == xval)
    matches <- as_tibble(as.data.frame(st_is_within_distance(coord_1, hex_centroids, 
                                                      dist = 2500))) %>% 
      select(col.id) %>% deframe()
    
    coord_matches <- hex_centroids_aspatial %>% 
      filter(hex_rowid %in% matches) %>% 
      select(hex_rowid, x, y) %>% 
      mutate(hex_rowid_1 = coord_1$hex_rowid,
             x_1 = coord_1$x,
             y_1 = coord_1$y) %>% 
      rename(hex_rowid_2 = hex_rowid,
             x_2 = x,
             y_2 = y) %>% 
      mutate(comb_id = paste0(as.character(hex_rowid_1), '_', as.character(hex_rowid_2)))
   
     distance_array_template <- rbind(distance_array_template, coord_matches)
  }
  return(distance_array_template)
}

# Florilegium: 

plan(callr, workers = 6)
options(future.globals.maxSize = 891289600)  

for  (i in 1:40) {
  ival = eval(i)
  edge_collection <- future_map(row_indices[ival], ~edge_generator(.x)) %>% bind_rows
  write_rds(edge_collection, paste0('outputs/network_components/edge_folder/edge_collection_',
                                   ival, '.rds'))
  print(paste('Completed:', ival))
}


# Advesperascit:

plan(callr, workers = 12)
options(future.globals.maxSize = 891289600)

for  (i in 40:80) {
  ival = eval(i)
  edge_collection <- future_map(row_indices[ival], ~edge_generator(.x)) %>% bind_rows
  write_rds(edge_collection, paste0('outputs/network_components/edge_folder/edge_collection_',
                                   ival, '.rds'))
  print(paste('Completed:', ival))
}


# Cognoscere

plan(callr, workers = 20)
options(future.globals.maxSize = 891289600)

for  (i in 150:160) {
  ival = eval(i)
  edge_collection <- future_map(row_indices[ival], ~edge_generator(.x)) %>% bind_rows
  write_rds(edge_collection, paste0('/home/steven/Downloads/edge_collection_',
                                   ival, '.rds'))
  print(paste('Completed:', ival))
}
  
####

edge_compilation <- list.files('outputs/network_components/edge_folder/', 
                            pattern = '*.rds',
                            full.names = TRUE,
                            recursive = TRUE)

edge_compilation <- lapply(edge_compilation, readRDS) %>% bind_rows()



edge_compilation <- edge_compilation %>% 
mutate(geometry = paste0('LINESTRING (', x_1, ' ', y_1, ', ', x_2, ' ', y_2, ')' ))

edge_compilation <- edge_compilation %>% st_as_sf(wkt = 'geometry', crs = 'ESRI:102008')


edge_compilation <-  edge_compilation %>% 
   rename(from = hex_rowid_1, to = hex_rowid_2)

edge_compilation$from <- as.character(edge_compilation$from)

edge_compilation$to <- as.character(edge_compilation$to)

duplicate_list <- st_equals(edge_compilation, retain_unique = TRUE)

edge_compilation <- edge_compilation[-unlist(duplicate_list),]

edge_compilation <- edge_compilation %>% mutate(eval_column = case_when(to == from ~ 'self', 
                                                                        to > from ~ 'not-self',
                                                                        to < from ~ 'not-self'))
edge_compilation <- edge_compilation %>% filter(eval_column == 'not-self') %>% 
  select(-eval_column)


edge_compilation <- edge_compilation %>% rowid_to_column('unique_id')

#write_sf(edge_compilation, 'outputs/network_components/edge_compilation_cleaned.gpkg')

edge_compilation <- read_sf('outputs/network_components/edge_compilation_cleaned.gpkg')

midpoint_generator <- function(row_subset) {
  
  midpoint_collection <- tibble(intermediate_id = numeric(),
                                to = character(),
                                x_2 = numeric(),
                                y_2 = numeric(),
                                from = character(),
                                x_1 = numeric(),
                                y_1 = numeric(),
                                comb_id = character(),
                                geom = st_sfc(crs = 'ESRI:102008'))
  
  for (i in row_subset) {
    ival = eval(i)
    int_filter <- edge_compilation %>% filter(unique_id == ival)
    int_midpoint <- suppressWarnings(st_centroid(int_filter)) %>%
      st_as_sf() %>% 
      mutate(intermediate_id = ival) %>% 
      select(-unique_id)
    midpoint_collection <- rbind(midpoint_collection, int_midpoint)
  }
  return(midpoint_collection)
  
}

# Cognoscere

plan(callr, workers = 20)
options(future.globals.maxSize = 891289600)

midpoint_collection <- future_map(1:250000, ~midpoint_generator(.x)) %>% bind_rows()

write_sf(midpoint_collection, 'outputs/network_components/midpoint_collection/midpoint_collection_1.gpkg')

# Advesperascit

plan(callr, workers = 12)
options(future.globals.maxSize = 891289600)

midpoint_collection <- future_map(250000:500000, ~midpoint_generator(.x)) %>% bind_rows()

write_sf(midpoint_collection, 'outputs/network_components/midpoint_collection/midpoint_collection_2.gpkg')

# Florilegium

plan(callr, workers = 8)
options(future.globals.maxSize = 891289600)

midpoint_collection <- future_map(500000:815410, ~midpoint_generator(.x)) %>% bind_rows()

write_sf(midpoint_collection, 'outputs/network_components/midpoint_collection/midpoint_collection_3.gpkg')

mc_1 <- read_sf('outputs/network_components/midpoint_collection/midpoint_collection_1.gpkg')
  
mc_2 <- read_sf('outputs/network_components/midpoint_collection/midpoint_collection_2.gpkg')
  
mc_3 <- read_sf('outputs/network_components/midpoint_collection/midpoint_collection_3.gpkg')


midpoint_combined <- bind_rows(mc_1, mc_2, mc_3)


mp_duplicate_list <- st_equals(midpoint_combined, retain_unique = TRUE)

midpoint_combined  <- midpoint_combined[-unlist(mp_duplicate_list),] 

midpoint_combined$hex_rowid <- 518770:1334179

midpoint_combined <- midpoint_combined %>% rowid_to_column('midpoint_id') %>% 
    mutate(midpoint_x = sf::st_coordinates(.)[,1],
         midpoint_y = sf::st_coordinates(.)[,2]) 

midpoint_aspatial <- midpoint_combined %>% st_drop_geometry()



final_edge_generator <- function(row_subset) {
  
  final_nodes_template <- tibble(midpoint_x = numeric(),
                      midpoint_y = numeric(),
                      orig_x = numeric(),
                      orig_y = numeric(),
                      midpoint_id = numeric(),
                      edge_pair_id = character()
                      )
  
  for (i in row_subset) {
  ival <- eval(i)
  int_midpoint <- midpoint_aspatial %>% filter(midpoint_id == ival)
  
  midpoint_1 <- tibble(midpoint_x = int_midpoint$midpoint_x,
                      midpoint_y = int_midpoint$midpoint_y,
                      orig_x = int_midpoint$x_2,
                      orig_y = int_midpoint$y_2,
                      orig_id = int_midpoint$to,
                      midpoint_id = int_midpoint$hex_rowid,
                      edge_pair_id = paste0(as.character(int_midpoint$to), '_', 
                                            as.character(int_midpoint$hex_rowid))
                      )
  
  midpoint_2 <- tibble(midpoint_x = int_midpoint$midpoint_x,
                      midpoint_y = int_midpoint$midpoint_y,
                      orig_x = int_midpoint$x_1,
                      orig_y = int_midpoint$y_1,
                      orig_id = int_midpoint$from,
                      midpoint_id = int_midpoint$hex_rowid,
                      edge_pair_id = paste0(as.character(int_midpoint$from), '_', 
                                            as.character(int_midpoint$hex_rowid))
                      )
  final_nodes_template <- bind_rows(final_nodes_template, midpoint_1, midpoint_2)
}

return(final_nodes_template)
  
}


plan(callr, workers = 20)
options(future.globals.maxSize = 891289600)

final_edge_collection <- future_map(1:815410, ~final_edge_generator(.x)) %>% bind_rows()

final_edge_collection_sf <- final_edge_collection %>% 
mutate(geometry = paste0('LINESTRING (', midpoint_x, ' ', midpoint_y, ', ', orig_x, ' ', orig_y, ')' ))

final_edge_collection_sf <- final_edge_collection_sf %>% st_as_sf(wkt = 'geometry', crs = 'ESRI:102008')

centroid_data <- hex_centroids_aspatial %>% select(hex_rowid, surface_type, lanes, speed_limit)

centroid_data$hex_rowid <- as.character(centroid_data$hex_rowid)

##

final_node_collection <- rbind(streamlined_centroid_data, midpoint_streamlined)

final_edge_collection_joined <- left_join(final_edge_collection_sf, 
                                          centroid_data, 
                                          by = c('orig_id' = 'hex_rowid') )


write_sf(final_edge_collection_joined, 'outputs/network_components/final_edge_collection.gpkg')

final_edge_collection_sfnetwork <- final_edge_collection_joined %>% 
  select(midpoint_id, orig_id, surface_type, lanes, speed_limit) %>% 
  rename(from = orig_id, to = midpoint_id)

write_sf(final_edge_collection_sfnetwork, 'outputs/network_components/edge_sfnetwork.gpkg')

###

midpoint_streamlined <- midpoint_combined %>% select('hex_rowid')

midpoint_streamlined$hex_rowid <- as.character(midpoint_streamlined$hex_rowid)

centroid_streamlined <- hex_centroids %>% select(hex_rowid)

centroid_streamlined$hex_rowid <- as.character(centroid_streamlined$hex_rowid)

final_node_collection <- bind_rows(centroid_streamlined, midpoint_streamlined)

write_sf(final_node_collection, 'outputs/network_components/final_node_collection.gpkg')

final_node_collection_sfnetwork <- final_node_collection %>% rename(index = hex_rowid)

write_sf(final_node_collection_sfnetwork, 'outputs/network_components/node_sfnetwork.gpkg')

```

# Point of origin and destination

```{r}

canada_metro_areas <- read_sf('processed/canada_metro_areas.gpkg') %>% st_transform(crs = 'ESRI:102008')

canada_filtered <- canada_metro_areas %>% select(CMANAME, PRNAME) %>%
  rename(metro_name = CMANAME, prov_state_name = PRNAME) %>% 
  mutate(country = 'Canada')

us_metro_areas <- read_sf('processed/us_metro_areas.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

us_filtered <- us_metro_areas %>% select(NAME)

hex_centroids <- read_sf('outputs/network_components/centroids.gpkg')


###

us_filtered <- us_filtered %>% separate(NAME, sep = ',', c('metro_name', 'prov_state_name'), remove = F) %>% select(metro_name, prov_state_name) %>% 
  mutate(country = 'US')

north_america_metros <- bind_rows(canada_filtered, us_filtered)

north_america_metros <- north_america_metros %>% rowid_to_column('metro_id')


origin_template <- tibble(x = numeric(),
                          y = numeric(),
                         metro_id = numeric(),
                         metro_name = character(),
                         prov_state_name = character(),
                         surface_type = character(),
                         country = character() )


for (i in 1:1094) {
  m_val <- eval(i)
  
  metro_subset <- north_america_metros %>% filter(metro_id == m_val)
  
  total_origins <- st_intersection(hex_centroids, metro_subset)
  
  metro_rows <- nrow(total_origins)
  
  if (metro_rows > 10) {
  origin_sample_size = 10
} else {
  origin_sample_size = metro_rows
}
  
  origin_sample <- slice_sample(total_origins, 
                                n = origin_sample_size) %>%
    st_drop_geometry()
  
  origin_template <- bind_rows(origin_template, origin_sample)
  
  print(paste('Completed:', i))
}


origin_sample_complete <- origin_template %>% select(x, y, metro_id, metro_name,
                                                     prov_state_name, country) %>% 
  st_as_sf(coords = c('x', 'y'), crs = 'ESRI:102008')

write_sf(origin_sample_complete, 'outputs/network_components/origin_collection.gpkg')

# Destination


dest_ids <- c(44296, 44199, 45796, 43087, 43163)
  
dest_names <- c('Kelowna', 'West Kelowna', 'Vernon', 'Summerland', 'Penticton')

destination_template <- tibble(id = numeric(),
                               name = character(),
                               geom = st_sfc(crs = 'ESRI:102008'))

for (i in 1:5) {
  ival = eval(i)
  dest_filtered <- final_node_collection_sfnetwork %>% filter(index == ival) %>% 
    select(-index) %>% 
    mutate(id = dest_ids[ival], name = dest_names[ival])
  destination_template <- bind_rows(destination_template, dest_filtered)
}

write_sf(destination_template, 'outputs/network_components/destination_collection.gpkg')
  
```

# Network traversal

```{r}

node_sfnetwork <- read_sf('outputs/network_components/node_sfnetwork.gpkg')

edge_sfnetwork <- read_sf('outputs/network_components/edge_sfnetwork.gpkg')

origins <- read_sf('outputs/network_components/origin_collection.gpkg')

destinations <- read_sf('outputs/network_components/destination_collection.gpkg')

###

origins_matched <- st_join(origins, node_sfnetwork) %>% 
  rowid_to_column('origin_id')

destinations_matched <- st_join(destinations, node_sfnetwork) %>% 
  rowid_to_column('destination_id')


network <- sfnetwork(node_sfnetwork, edge_sfnetwork,
                     directed = FALSE,
                     node_key = 'index',
                     force = TRUE) %>% 
   activate("edges") %>%
  mutate(weight = edge_length())


path_collection_template <- tibble(path_id = numeric(),
                                   origin = character(),
                          destination = character(),
                          geom = st_sfc(crs = 'ESRI:102008'))

edge_sfnetwork$to <- as.character(edge_sfnetwork$to)
edge_sfnetwork$from <- as.character(edge_sfnetwork$from)


router <- function(row_subset) {
  for (i in row_subset) {
  ival = eval(i)
  int_origin <- origins_matched %>% filter(origin_id == ival)
  int_origin_index <- int_origin$index
  int_path <- suppressWarnings(st_network_paths(network, 
                                                from = int_origin_index, 
                                                to = 43087, 
                                                weights = 'weight'))
  if (length(unlist(int_path$node_paths)) > 0) {
    int_nodes <- as_tibble(unlist(int_path$node_paths)) %>% 
      rename(index = value)
    int_nodes$index <- as.character(int_nodes$index)
    int_node_spatial <- inner_join(node_sfnetwork, int_nodes, by = 'index')
    test_edge_1 <- inner_join(edge_sfnetwork, int_nodes, by = c('to' = 'index'))
    path_edges <- inner_join(test_edge_1, int_nodes, by = c('from' = 'index'))
    path_edges <- path_edges %>% select(geom)
    path_edges <- st_union(path_edges)
    path_collection_int <- tibble(path_id = ival,
                                origin = int_origin$metro_name,
                          destination = 'Summerland',
                          geom = path_edges)
    path_collection_template <- bind_rows(path_collection_template, path_collection_int)
  } 
}
  return(path_collection_template)
}


plan(callr, workers = 6)
options(future.globals.maxSize = 1.6e+10)

route_collection <- future_map(1:10642, ~router(.x)) %>% bind_rows() %>% st_as_sf()

write_sf(route_collection, 'outputs/network_components/summerland_route_collection.gpkg')





```

# Network routing

```{r}


```


