# Zebra mussel network modeling script

# Contact information

Contact: Steven Brownlee
Email: steven.fr.brownlee@gmail.com
Date last revised: June 24 2024

# Session info

R version 4.4.1 (2024-06-14)
Platform: x86_64-redhat-linux-gnu
Running under: Nobara Linux 40 (GNOME Edition)

Matrix products: default
BLAS/LAPACK: FlexiBLAS OPENBLAS-OPENMP;  LAPACK version 3.11.0

locale:
 [1] LC_CTYPE=en_CA.UTF-8      LC_NUMERIC=C              LC_TIME=en_CA.utf8       
 [4] LC_COLLATE=en_CA.UTF-8    LC_MONETARY=en_CA.utf8    LC_MESSAGES=en_CA.UTF-8  
 [7] LC_PAPER=en_CA.utf8       LC_NAME=C                 LC_ADDRESS=C             
[10] LC_TELEPHONE=C            LC_MEASUREMENT=en_CA.utf8 LC_IDENTIFICATION=C      

time zone: America/Vancouver
tzcode source: system (glibc)

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

loaded via a namespace (and not attached):
 [1] digest_0.6.36     tidyr_1.3.1       utf8_1.2.4        R6_2.5.1         
 [5] fastmap_1.2.0     tidyselect_1.2.1  xfun_0.46         magrittr_2.0.3   
 [9] glue_1.7.0        tibble_3.2.1      knitr_1.48        pkgconfig_2.0.3  
[13] htmltools_0.5.8.1 rmarkdown_2.27    dplyr_1.1.4       generics_0.1.3   
[17] lifecycle_1.0.4   cli_3.6.3         fansi_1.0.6       vctrs_0.6.5      
[21] compiler_4.4.1    purrr_1.0.2       rstudioapi_0.16.0 tools_4.4.1      
[25] pillar_1.9.0      evaluate_0.24.0   yaml_2.3.10       rlang_1.1.4   

# Set up data directory

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/home/steven/Documents/workspace/thesis/ch2/gis_data/')
```

# R library setup

```{r}
library(sf)
library(tidyverse)
library(rcartocolor)
library(sfnetworks)
library(qgisprocess)
library(future)
library(future.callr)
library(purrr)
library(furrr)
library(progressr)
library(stars)
library(paletteer)
library(ggnewscale)
library(terra)
```

# Import shapefiles

```{r}
zm_occurrences <- read_sf('processed/dreissenid_occurrences.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

us_metro <- read_sf('processed/us_metro_areas.gpkg')%>% 
  st_transform(crs = 'ESRI:102008')

canada_metro <- read_sf('processed/canada_metro_areas.gpkg')%>% 
  st_transform(crs = 'ESRI:102008')

land <- read_sf('processed/shared_north_america_coastline.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

lakes <- read_sf('processed/cec_lakes.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

usdot_road_network <- read_sf('processed/usdot_road_network.gpkg')%>% 
  st_transform(crs = 'ESRI:102008')

qgis_bbox <- read_sf('processed/qgis_bbox.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

```


# Pre-processing

```{r}
# Create buffers around ZM occurrences

zm_10 <- st_buffer(zm_occurrences, dist = 10000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = 10)

zm_20 <- st_buffer(zm_occurrences, dist = 20000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = 20)

zm_30 <- st_buffer(zm_occurrences, dist = 30000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = 30)

zm_40 <- st_buffer(zm_occurrences, dist = 40000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = 40)

zm_50 <- st_buffer(zm_occurrences, dist = 50000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = 50)

zm_buffer_collected <- rbind(zm_50, zm_40, zm_30, zm_20, zm_10)

rm(zm_50, zm_40, zm_30, zm_20, zm_10)

write_sf(zm_buffer_collected, 'processed/zm_buffer.gpkg')

# Merge US and Canada metro areas

us_metro <- us_metro %>% select(NAME) %>% rename(metro_area = NAME)

canada_metro <- canada_metro %>% select(CMANAME) %>% rename(metro_area = CMANAME)

north_america_metro <- rbind(us_metro, canada_metro)

# Create node network for USDOT roads
```

# Split road into manageable chunks

```{r}

usdot_road_network <- usdot_road_network %>% rownames_to_column('seg_id')


plan(callr, workers = 6)
options(future.globals.maxSize= 891289600)

node_generator <- function(sublist) {

for (x in sublist) {
  usdot_template <- tibble(x_coord = numeric(),
                         y_coord = numeric(),
                         country = character(),
                         usdot_id = numeric(),
                         jurisdiction = character(),
                         road_number = character(),
                         road_name = character(),
                         surface_type = character(),
                         lanes = numeric(),
                         speed_limit = numeric(),
                         class = numeric())
  
segment_filtered <- usdot_road_network %>% filter(seg_id == x)
length_extract = segment_filtered$LENGTH

if (length_extract <= 1) {
  sample_size = 5
} else if (1 < length_extract & length_extract <= 5 ) {
  sample_size = 20
} else if (5 < length_extract & length_extract <= 20) {
  sample_size = 100 
} else if (20 < length_extract & length_extract <= 50) {
  sample_size = 400 
} else if (50 < length_extract & length_extract <= 100) {
  sample_size = 700
} else if (100 < length_extract & length_extract <= 400) {
  sample_size = 2500
} else {sample_size = 2000} 

usdot_nodes <- st_sample(segment_filtered, size = sample_size, type = 'regular') %>% 
  st_cast("POINT") %>% 
  st_as_sf() %>% 
  mutate(x_coord = sf::st_coordinates(.)[,1],
         y_coord = sf::st_coordinates(.)[,2],
         usdot_id = segment_filtered$ID,
         country = segment_filtered$COUNTRY,
         jurisdiction = segment_filtered$JURISNAME,
         road_number = segment_filtered$ROADNUM,
         road_name = segment_filtered$ROADNAME,
         surface_type = segment_filtered$SURFACE,
         lanes = segment_filtered$LANES,
         speed_limit = segment_filtered$SPEEDLIM,
         class = segment_filtered$CLASS) %>% 
  st_drop_geometry()

usdot_nodes$x_coord <- as.numeric(usdot_nodes$x_coord)
usdot_nodes$y_coord <- as.numeric(usdot_nodes$y_coord)

usdot_template <- rbind(usdot_template, usdot_nodes) 

}
usdot_filled <- usdot_template
return(usdot_filled)
}

usdot_nodes <- future_map(1:716638, ~node_generator(.x))

###

usdot_nodes_2 <- bind_rows(usdot_nodes)

write_rds(usdot_nodes_2, 'outputs/node_folder/usdot_combined_2.rds')

rm(usdot_nodes_2, usdot_nodes)

usdot_combined <- read_rds('outputs/node_folder/usdot_combined_2.rds') 

usdot_combined <- usdot_combined %>% rowid_to_column('node_id')


inside_list = c(1, 500000,
                1000000,
                2000000,
                3000000,
                4000000,
                5000000,
                6000000,
                7000000,
                8000000,
                9000000,
                10000000,
                11000000)

outside_list= c(500000,
                1000000,
                2000000,
                3000000,
                4000000,
                5000000,
                6000000,
                7000000,
                8000000,
                9000000,
                10000000,
                11000000,
                11544435)

for (i in 1:13) {
  inside_index <- inside_list[i]
  outside_index <- outside_list[i]
  usdot_subset <- usdot_combined %>% 
    filter(node_id >= inside_index & node_id < outside_index)
   usdot_subset <- usdot_subset %>% 
     drop_na() %>% 
  st_as_sf(coords = c('x_coord', 'y_coord'), crs = st_crs('ESRI:102008'))
  write_sf(usdot_subset, paste0('outputs/node_folder/node_pieces/node_subset_', i, '.gpkg'))
  print(paste('Completed:', i))
}

###
```


# Join road data to grid
```{r}
#overlay_grid <- st_make_grid(qgis_bbox, cellsize = 2000,
#                             what = 'polygons', square = FALSE) %>% st_as_sf() %>% 
#  st_transform(crs = 'ESRI:102008')

#write_sf(overlay_grid, 'processed/2km_hexagon_grid.gpkg')

overlay_grid <- read_sf('processed/2km_hexagon_grid.gpkg')

overlay_grid <- overlay_grid %>% rowid_to_column('hex_id')

overlay_grid$hex_group <- as.numeric(cut_number(overlay_grid$hex_id, 1000))

for (x in 1:1000) {
  overlay_subset <- overlay_grid %>% filter(hex_group == x)
  write_sf(overlay_subset, paste0('outputs/overlay_grid_segments/hex_group_', x, '.gpkg'))
  print(paste('Completed:', x))
}


hex_list <- list.files('outputs/overlay_grid_segments/', 
                            pattern = '*.gpkg',
                            full.names = TRUE,
                            recursive = TRUE)

node_list <- list.files('outputs/node_folder/node_pieces/', 
                            pattern = '*.gpkg',
                            full.names = TRUE,
                            recursive = TRUE)


table_template <- tibble(hexid = numeric(),
                         node_id = numeric(),
                         usdot_id = numeric(),
                         country = numeric(),
                         jurisdiction = character(),
                         road_number = character(),
                         road_name = character(),
                         surface_type = character(),
                         lanes = numeric())

joiner <- function(sequence) {
for (x in sequence) {
  hex_segment  <- read_sf(hex_list[x]) %>% select(-hex_group)
  for (i in 1:13) {
    interior_nodes  <- read_sf(node_list[i])
    hex_interior <- st_join(hex_segment, interior_nodes) %>% st_drop_geometry()
    table_template <- rbind(table_template, hex_interior)
    print(paste('Interior done:', i))
}
  table_export <- table_template
  write_rds(table_export, paste0('outputs/overlay_grid_joined/overlay_grid_joined_', x, '.rds'))
  table_template <- tibble(hexid = numeric(),
                         node_id = numeric(),
                         usdot_id = numeric(),
                         country = numeric(),
                         jurisdiction = character(),
                         road_number = character(),
                         road_name = character(),
                         surface_type = character(),
                         lanes = numeric())
  return(table_export)
}
}

plan(callr, workers = 6)
options(future.globals.maxSize= 891289600)


joined <- future_map(1:1000, ~joiner(.x))

joined <- bind_rows(joined)

write_rds(joined, 'outputs/overlay_grid_joined/overlay_tables/joined_table.rds')

joined_table <- drop_na(joined)

write_rds(joined_table, 'outputs/overlay_grid_joined/overlay_tables/joined_table_na_dropped.rds')

joined_table <- read_rds('outputs/overlay_grid_joined/overlay_tables/joined_table_na_dropped.rds')

joined_table_summarized <- joined_table %>% select(hex_id, country, jurisdiction,
                                                   surface_type, surface_type,
                                                   lanes, speed_limit, class)

joined_table_summarized$surface_type <- case_match(joined_table_summarized$surface_type, 
                                            'Unpaved' ~ '1', 
                                            'Paved' ~ '2')

joined_table_summarized$surface_type <- as.numeric(joined_table_summarized$surface_type)

joined_table_summarized <- joined_table %>% 
  group_by(hex_id) %>% 
  summarize(surface_type = max(surface_type),
            lanes = max(lanes),
            speed_limit = max(speed_limit))

hex_final <- left_join(overlay_grid, joined_table_summarized, by = 'hex_id') %>% 
  drop_na()


write_sf(hex_final, 'outputs/network_components/hex_grid_final.gpkg')

```

# Edge generation

```{r}
hexagon_grid <- read_sf('outputs/network_components/hex_grid_final.gpkg') %>% 
  rowid_to_column('hex_rowid')

hex_centroids <- hexagon_grid %>% st_centroid() %>%  
  mutate(x = sf::st_coordinates(.)[,1],
         y = sf::st_coordinates(.)[,2]) %>% 
  select(hex_rowid, surface_type, lanes, speed_limit, x, y, geom)

#write_sf(hex_centroids, 'outputs/network_components/centroids.gpkg')

hex_centroids_aspatial <- hex_centroids %>% st_drop_geometry()

###

row_indices <- split(hex_centroids$hex_rowid, cut(hex_centroids$hex_rowid, breaks = 160))


###

edge_generator <- function(row_subset) {
  
  distance_array_template <- tibble(hex_rowid_1 = numeric(),
                                    x_1 = numeric(),
                                    y_1 = numeric(),
                                    hex_rowid_2 = numeric(),
                                    x_2 = numeric(),
                                    y_2 = numeric(),
                                    comb_hex_rowid = character())
  
  for (x in row_subset) {
    xval = eval(x)
    coord_1 <- hex_centroids %>% filter(hex_rowid == xval)
    matches <- as_tibble(as.data.frame(st_is_within_distance(coord_1, hex_centroids, 
                                                      dist = 2500))) %>% 
      select(col.id) %>% deframe()
    
    coord_matches <- hex_centroids_aspatial %>% 
      filter(hex_rowid %in% matches) %>% 
      select(hex_rowid, x, y) %>% 
      mutate(hex_rowid_1 = coord_1$hex_rowid,
             x_1 = coord_1$x,
             y_1 = coord_1$y) %>% 
      rename(hex_rowid_2 = hex_rowid,
             x_2 = x,
             y_2 = y) %>% 
      mutate(comb_id = paste0(as.character(hex_rowid_1), '_', as.character(hex_rowid_2)))
   
     distance_array_template <- rbind(distance_array_template, coord_matches)
  }
  return(distance_array_template)
}

# Florilegium: 

plan(callr, workers = 6)
options(future.globals.maxSize = 891289600)  

for  (i in 1:40) {
  ival = eval(i)
  edge_collection <- future_map(row_indices[ival], ~edge_generator(.x)) %>% bind_rows
  write_rds(edge_collection, paste0('outputs/network_components/edge_folder/edge_collection_',
                                   ival, '.rds'))
  print(paste('Completed:', ival))
}


# Advesperascit:

plan(callr, workers = 12)
options(future.globals.maxSize = 891289600)

for  (i in 40:80) {
  ival = eval(i)
  edge_collection <- future_map(row_indices[ival], ~edge_generator(.x)) %>% bind_rows
  write_rds(edge_collection, paste0('outputs/network_components/edge_folder/edge_collection_',
                                   ival, '.rds'))
  print(paste('Completed:', ival))
}


# Cognoscere

plan(callr, workers = 20)
options(future.globals.maxSize = 891289600)

for  (i in 150:160) {
  ival = eval(i)
  edge_collection <- future_map(row_indices[ival], ~edge_generator(.x)) %>% bind_rows
  write_rds(edge_collection, paste0('/home/steven/Downloads/edge_collection_',
                                   ival, '.rds')) 
  
  print(paste('Completed:', ival))
}
  
####

edge_compilation <- list.files('outputs/network_components/edge_folder/', 
                            pattern = '*.rds',
                            full.names = TRUE,
                            recursive = TRUE)

edge_compilation <- lapply(edge_compilation, readRDS) %>% bind_rows()



edge_compilation <- edge_compilation %>% 
mutate(geometry = paste0('LINESTRING (', x_1, ' ', y_1, ', ', x_2, ' ', y_2, ')' ))

edge_compilation <- edge_compilation %>% st_as_sf(wkt = 'geometry', crs = 'ESRI:102008')


edge_compilation <-  edge_compilation %>% 
   rename(from = hex_rowid_1, to = hex_rowid_2)

edge_compilation$from <- as.character(edge_compilation$from)

edge_compilation$to <- as.character(edge_compilation$to)

duplicate_list <- st_equals(edge_compilation, retain_unique = TRUE)

edge_compilation <- edge_compilation[-unlist(duplicate_list),]

edge_compilation <- edge_compilation %>% mutate(eval_column = case_when(to == from ~ 'self', 
                                                                        to > from ~ 'not-self',
                                                                        to < from ~ 'not-self'))
edge_compilation <- edge_compilation %>% filter(eval_column == 'not-self') %>% 
  select(-eval_column)


edge_compilation <- edge_compilation %>% rowid_to_column('unique_id')

#write_sf(edge_compilation, 'outputs/network_components/edge_compilation_cleaned.gpkg')

edge_compilation <- read_sf('outputs/network_components/edge_compilation_cleaned.gpkg')

midpoint_generator <- function(row_subset) {
  
  midpoint_collection <- tibble(intermediate_id = numeric(),
                                to = character(),
                                x_2 = numeric(),
                                y_2 = numeric(),
                                from = character(),
                                x_1 = numeric(),
                                y_1 = numeric(),
                                comb_id = character(),
                                geom = st_sfc(crs = 'ESRI:102008'))
  
  for (i in row_subset) {
    ival = eval(i)
    int_filter <- edge_compilation %>% filter(unique_id == ival)
    int_midpoint <- suppressWarnings(st_centroid(int_filter)) %>%
      st_as_sf() %>% 
      mutate(intermediate_id = ival) %>% 
      select(-unique_id)
    midpoint_collection <- rbind(midpoint_collection, int_midpoint)
  }
  return(midpoint_collection)
  
}

# Cognoscere

plan(callr, workers = 20)
options(future.globals.maxSize = 891289600)

midpoint_collection <- future_map(1:250000, ~midpoint_generator(.x)) %>% bind_rows()

write_sf(midpoint_collection, 'outputs/network_components/midpoint_collection/midpoint_collection_1.gpkg')

# Advesperascit

plan(callr, workers = 12)
options(future.globals.maxSize = 891289600)

midpoint_collection <- future_map(250000:500000, ~midpoint_generator(.x)) %>% bind_rows()

write_sf(midpoint_collection, 'outputs/network_components/midpoint_collection/midpoint_collection_2.gpkg')

# Florilegium

plan(callr, workers = 8)
options(future.globals.maxSize = 891289600)

midpoint_collection <- future_map(500000:815410, ~midpoint_generator(.x)) %>% bind_rows()

write_sf(midpoint_collection, 'outputs/network_components/midpoint_collection/midpoint_collection_3.gpkg')

mc_1 <- read_sf('outputs/network_components/midpoint_collection/midpoint_collection_1.gpkg')
  
mc_2 <- read_sf('outputs/network_components/midpoint_collection/midpoint_collection_2.gpkg')
  
mc_3 <- read_sf('outputs/network_components/midpoint_collection/midpoint_collection_3.gpkg')


midpoint_combined <- bind_rows(mc_1, mc_2, mc_3)


mp_duplicate_list <- st_equals(midpoint_combined, retain_unique = TRUE)

midpoint_combined  <- midpoint_combined[-unlist(mp_duplicate_list),] 

midpoint_combined$hex_rowid <- 518770:1334179

midpoint_combined <- midpoint_combined %>% rowid_to_column('midpoint_id') %>% 
    mutate(midpoint_x = sf::st_coordinates(.)[,1],
         midpoint_y = sf::st_coordinates(.)[,2]) 

midpoint_aspatial <- midpoint_combined %>% st_drop_geometry()



final_edge_generator <- function(row_subset) {
  
  final_nodes_template <- tibble(midpoint_x = numeric(),
                      midpoint_y = numeric(),
                      orig_x = numeric(),
                      orig_y = numeric(),
                      midpoint_id = numeric(),
                      edge_pair_id = character()
                      )
  
  for (i in row_subset) {
  ival <- eval(i)
  int_midpoint <- midpoint_aspatial %>% filter(midpoint_id == ival)
  
  midpoint_1 <- tibble(midpoint_x = int_midpoint$midpoint_x,
                      midpoint_y = int_midpoint$midpoint_y,
                      orig_x = int_midpoint$x_2,
                      orig_y = int_midpoint$y_2,
                      orig_id = int_midpoint$to,
                      midpoint_id = int_midpoint$hex_rowid,
                      edge_pair_id = paste0(as.character(int_midpoint$to), '_', 
                                            as.character(int_midpoint$hex_rowid))
                      )
  
  midpoint_2 <- tibble(midpoint_x = int_midpoint$midpoint_x,
                      midpoint_y = int_midpoint$midpoint_y,
                      orig_x = int_midpoint$x_1,
                      orig_y = int_midpoint$y_1,
                      orig_id = int_midpoint$from,
                      midpoint_id = int_midpoint$hex_rowid,
                      edge_pair_id = paste0(as.character(int_midpoint$from), '_', 
                                            as.character(int_midpoint$hex_rowid))
                      )
  final_nodes_template <- bind_rows(final_nodes_template, midpoint_1, midpoint_2)
}

return(final_nodes_template)
  
}


plan(callr, workers = 20)
options(future.globals.maxSize = 891289600)

final_edge_collection <- future_map(1:815410, ~final_edge_generator(.x)) %>% bind_rows()

final_edge_collection_sf <- final_edge_collection %>% 
mutate(geometry = paste0('LINESTRING (', midpoint_x, ' ', midpoint_y, ', ', orig_x, ' ', orig_y, ')' ))

final_edge_collection_sf <- final_edge_collection_sf %>% st_as_sf(wkt = 'geometry', crs = 'ESRI:102008')

centroid_data <- hex_centroids_aspatial %>% select(hex_rowid, surface_type, lanes, speed_limit)

centroid_data$hex_rowid <- as.character(centroid_data$hex_rowid)

##

final_node_collection <- rbind(streamlined_centroid_data, midpoint_streamlined)

final_edge_collection_joined <- left_join(final_edge_collection_sf, 
                                          centroid_data, 
                                          by = c('orig_id' = 'hex_rowid') )


write_sf(final_edge_collection_joined, 'outputs/network_components/final_edge_collection.gpkg')

final_edge_collection_sfnetwork <- final_edge_collection_joined %>% 
  select(midpoint_id, orig_id, surface_type, lanes, speed_limit) %>% 
  rename(from = orig_id, to = midpoint_id)

write_sf(final_edge_collection_sfnetwork, 'outputs/network_components/edge_sfnetwork.gpkg')

###

midpoint_streamlined <- midpoint_combined %>% select('hex_rowid')

midpoint_streamlined$hex_rowid <- as.character(midpoint_streamlined$hex_rowid)

centroid_streamlined <- hex_centroids %>% select(hex_rowid)

centroid_streamlined$hex_rowid <- as.character(centroid_streamlined$hex_rowid)

final_node_collection <- bind_rows(centroid_streamlined, midpoint_streamlined)

write_sf(final_node_collection, 'outputs/network_components/final_node_collection.gpkg')

final_node_collection_sfnetwork <- final_node_collection %>% rename(index = hex_rowid)

write_sf(final_node_collection_sfnetwork, 'outputs/network_components/node_sfnetwork.gpkg')

```

# Point of origin and destination

```{r}

canada_metro_areas <- read_sf('processed/canada_metro_areas.gpkg') %>% st_transform(crs = 'ESRI:102008')

canada_filtered <- canada_metro_areas %>% select(CMANAME, PRNAME) %>%
  rename(metro_name = CMANAME, prov_state_name = PRNAME) %>% 
  mutate(country = 'Canada')

us_metro_areas <- read_sf('processed/us_metro_areas.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

us_filtered <- us_metro_areas %>% select(NAME)

hex_centroids <- read_sf('outputs/network_components/centroids.gpkg')


###

us_filtered <- us_filtered %>% separate(NAME, sep = ',', c('metro_name', 'prov_state_name'), remove = F) %>% select(metro_name, prov_state_name) %>% 
  mutate(country = 'US')

north_america_metros <- bind_rows(canada_filtered, us_filtered)

north_america_metros <- north_america_metros %>% rowid_to_column('metro_id')

```

# Points of origin

```{r}


origin_template <- tibble(x = numeric(),
                          y = numeric(),
                         metro_id = numeric(),
                         metro_name = character(),
                         prov_state_name = character(),
                         country = character() )


for (i in 1:1094) {
  m_val <- eval(i)
  
  metro_subset <- north_america_metros %>% filter(metro_id == m_val)
  
  total_origins <- st_intersection(hex_centroids, metro_subset)
  
  metro_rows <- nrow(total_origins)
  
  if (metro_rows > 10) {
  origin_sample_size = 10
} else {
  origin_sample_size = metro_rows
}
  
  origin_sample <- slice_sample(total_origins, 
                                n = origin_sample_size) %>%
    st_drop_geometry()
  
  origin_template <- bind_rows(origin_template, origin_sample)
  
  print(paste('Completed:', i))
}


origin_sample_complete <- origin_template %>% select(x, y, metro_id, metro_name,
                                                     prov_state_name, country) %>% 
  st_as_sf(coords = c('x', 'y'), crs = 'ESRI:102008')

write_sf(origin_sample_complete, 'outputs/network_components/origin_collection.gpkg')

```

# Destinations, metro areas

```{r}

# Destination


na_political_bound <- read_sf('downloads/political_boundaries/boundaries_p_2021_v3.shp') %>% 
  st_transform(crs = 'ESRI:102008')

bc_bound <- na_political_bound %>% 
  filter(NAME_En == 'British Columbia')

destination_list <- st_intersection(canada_metro_areas, bc_bound)

destination_list <- destination_list %>% rename(metro_name = CMANAME) %>% 
  select(metro_name) %>% 
  filter(metro_name != 'Whitehorse') %>% 
  rowid_to_column('destination_id')

##

destination_template <- tibble(x = numeric(),
                          y = numeric(),
                         destination_name = character(),
                         geom = st_sfc(crs = 'ESRI:102008')
                         ) %>% st_as_sf()


for (i in 1:26) {
  d_val <- eval(i)
  
  destination_subset <- destination_list %>% filter(destination_id == d_val)
  
  destination_area <- as.numeric(st_area(destination_subset)/1000000)
  
  if (destination_area > 10) {
  destination_sample_size = 10
} else {
  destination_sample_size = 5
}
 
  destination_sample <- st_sample(destination_subset, 
                                  size = destination_sample_size) %>% 
  st_cast("POINT") %>% 
  st_as_sf() %>% 
  rename(geom = x) %>% 
  mutate(x = sf::st_coordinates(.)[,1],
         y = sf::st_coordinates(.)[,2],
         destination_name = destination_subset$metro_name
         ) 
  destination_template <- bind_rows(destination_template, destination_sample)
  
  print(paste('Completed:', i))
}

destination_template <- destination_template %>% rowid_to_column('destination_id')


write_sf(destination_template, 'outputs/network_components/destinations/metro/metro_destinations.gpkg')
  

```

# Destinations, lakes

```{r}
cci_lakes <- read_sf('processed/cci_lakes.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

cci_lakes_subset <- cci_lakes %>% filter(id == 300000651 | id == 384 | 
                                           id == 300008244| id == 300000579 |
                                           id == 300006584 | id == 565 |
                                           id == 501
                                         )

name_corrector <- tibble(id = c(384, 501, 565, 
                                300000579, 300000651, 300006584,
                                300008244),
                         new_name = c('Babine', 'Stuart', 'Okanagan',
                                      'Francois', 'Adams', 'Fraser',
                                      'Cowichan'))

cci_lakes_subset <- left_join(cci_lakes_subset, name_corrector, by = 'id') %>% 
  select(new_name) %>% 
  rename(lake_name = new_name)

write_sf(cci_lakes_subset, 'processed/cci_lakes_sample.gpkg')

cci_buffered_big <- st_buffer(cci_lakes_subset, dist = 5000)

cci_buffered_small <- st_buffer(cci_lakes_subset, dist = 0.01)

st_erase = function(x, y) st_difference(x, st_union(st_combine(y)))

lake_destinations <- st_erase(cci_buffered_big, cci_buffered_small)

lake_destinations <- lake_destinations %>% rowid_to_column('lake_id')

###

lake_destination_template <- tibble(x = numeric(),
                          y = numeric(),
                         destination_name = character(),
                         geom = st_sfc(crs = 'ESRI:102008')
                         ) %>% st_as_sf()

for (i in 1:7) {
   d_val <- eval(i)
   destination_subset <- lake_destinations %>% filter(lake_id == d_val)
   destination_sample <- st_sample(destination_subset, 
                                  size = 10) %>% 
     st_cast("POINT") %>% 
     st_as_sf() %>% 
     rename(geom = x) %>% mutate(x = sf::st_coordinates(.)[,1],
         y = sf::st_coordinates(.)[,2],
         destination_name = destination_subset$lake_name
         ) 
     lake_destination_template <- bind_rows(lake_destination_template, destination_sample)
  
  print(paste('Completed:', i))
}

lake_destination_template <- lake_destination_template %>% 
  rowid_to_column('destination_id')

write_sf(lake_destination_template, 'outputs/network_components/destinations/lake/lake_destinations.gpkg')

```

# Destinations, Okanagan

```{r}

cci_lakes_subset <- read_sf('processed/cci_lakes_sample.gpkg')

okanagan <- cci_lakes_subset %>% filter(lake_name == 'Okanagan')

okanagan_buffered_big <- st_buffer(okanagan, dist = 5000)

okanagan_buffered_small <- st_buffer(okanagan, dist = 0.01)

st_erase = function(x, y) st_difference(x, st_union(st_combine(y)))

okanagan_template <- st_erase(okanagan_buffered_big, okanagan_buffered_small)

okanagan_samples <- st_sample(okanagan_template, 
                                  size = 20) %>% 
     st_cast("POINT") %>% 
     st_as_sf() %>% 
     rename(geom = x) %>% mutate(x = sf::st_coordinates(.)[,1],
         y = sf::st_coordinates(.)[,2],
         destination_name = 'Okanagan'
         ) 


okanagan_samples <- okanagan_samples %>% 
  rowid_to_column('destination_id')

write_sf(okanagan_samples, 'outputs/network_components/destinations/okanagan/okanagan_destinations.gpkg')

```


# Network traversal

```{r}

node_sfnetwork <- read_sf('outputs/network_components/node_sfnetwork.gpkg')

edge_sfnetwork <- read_sf('outputs/network_components/edge_sfnetwork.gpkg')

origins <- read_sf('outputs/network_components/origin_collection.gpkg') %>% 
  rowid_to_column('origin_id')

origins_matched <- st_join(origins, node_sfnetwork) 

write_sf(origins_matched, 'outputs/network_components/origin_matched.gpkg')

###



network <- sfnetwork(node_sfnetwork, edge_sfnetwork,
                     directed = FALSE,
                     node_key = 'index',
                     force = TRUE) %>% 
   activate("edges") %>%
  mutate(weight = edge_length())

edge_sfnetwork$to <- as.character(edge_sfnetwork$to)
edge_sfnetwork$from <- as.character(edge_sfnetwork$from)

write_sf(edge_sfnetwork, 'outputs/network_components/edge_sfnetwork_routing.gpkg')

write_rds(network, 'outputs/network_components/sf_network.rds')

###

```

# Route point function

```{r}

router <- function(row_subset, destination) {
  
  outer_template_table <- tibble(index = character(),
                           destination_id = numeric(),
                           origin_id = numeric(),
                           node_position = numeric(),
                           x = numeric(),
                           y = numeric())
  
  dest_row <- nrow(destination)
  
  for (i in row_subset) {
    
      int_origin <- origins_matched %>% filter(origin_id == i)
    
      inner_template_table <- tibble(index = character(),
                           destination_id = numeric(),
                           origin_id = numeric(),
                           node_position = numeric(),
                           x = numeric(),
                           y = numeric())
     
     int_path <- suppressWarnings(st_network_paths(network, 
                                                from = int_origin, 
                                                to = destination, 
                                                weights = 'weight'))
    
     for (j in 1:dest_row) {
           int_path_tibble <- as_tibble(unlist(int_path[j,1])) %>% 
             rename(index = value) %>% 
             mutate(destination_id = j) %>% 
             mutate(origin_id = int_origin$origin_id) %>% 
             rowid_to_column('node_position')
           int_path_tibble$index <- as.character(int_path_tibble$index)
           suppressMessages(int_path_joined <- left_join(int_path_tibble, node_aspatial))
           inner_template_table <- bind_rows(inner_template_table, int_path_joined)
     }

    outer_template_table <- bind_rows(outer_template_table, inner_template_table)
  }
  
  return(outer_template_table)
  
}

  
```



# Route generation, metros

```{r}

metro_dest <- read_sf('outputs/network_components/destinations/metro/metro_destinations.gpkg')

origins_matched <- read_sf('outputs/network_components/origin_matched.gpkg')

network <- readRDS('outputs/network_components/sf_network.rds')

node_sfnetwork <- read_sf('outputs/network_components/node_sfnetwork.gpkg')

node_aspatial <- node_sfnetwork %>% mutate(x = sf::st_coordinates(.)[,1],
         y = sf::st_coordinates(.)[,2],) %>% 
  st_drop_geometry()



###

plan(callr, workers = 6)
options(future.globals.maxSize = 1.6e+10)

groups <- split(1:10612, ceiling(seq_along(1:10612)/25))

###

for (x in 1:length(groups)) {
  int_vector <- unlist(groups[x])
  
  metro_table <- future_map(int_vector, ~router(.x, destination = metro_dest)) %>% 
  bind_rows()
  
  metro_path <- paste('outputs/network_components/routes/metro/metro_route_', x, '.rds')
  
  write_rds(metro_table, metro_path)
  
  print(paste('Completed:', x))
}





```


# Route generation, lakes

```{r}

lake_dest <- read_sf('outputs/network_components/destinations/lake/lake_destinations.gpkg')

origins_matched <- read_sf('outputs/network_components/origin_matched.gpkg')

network <- readRDS('outputs/network_components/sf_network.rds')

node_sfnetwork <- read_sf('outputs/network_components/node_sfnetwork.gpkg')

node_aspatial <- node_sfnetwork %>% mutate(x = sf::st_coordinates(.)[,1],
         y = sf::st_coordinates(.)[,2],) %>% 
  st_drop_geometry()

###

plan(callr, workers = 6)
options(future.globals.maxSize = 1.6e+10)

groups <- split(1:10612, ceiling(seq_along(1:10612)/25))

###


for (x in 1:length(groups)) {
  
  int_vector <- unlist(groups[x])
  
  lake_table <- future_map(int_vector, ~router(.x, destination = lake_dest)) %>% 
  bind_rows()
  
  lake_path <- paste0('outputs/network_components/routes/lake/lake_route_', x, '.rds')
  
  write_rds(lake_table, lake_path)
  print(paste('Completed:', x))
}


```

# Route generation, Okanagan

```{r}

ok_dest <- read_sf('outputs/network_components/destinations/okanagan/okanagan_destinations.gpkg')

origins_matched <- read_sf('outputs/network_components/origin_matched.gpkg')

network <- readRDS('outputs/network_components/sf_network.rds')

node_sfnetwork <- read_sf('outputs/network_components/node_sfnetwork.gpkg')

node_aspatial <- node_sfnetwork %>% mutate(x = sf::st_coordinates(.)[,1],
         y = sf::st_coordinates(.)[,2],) %>% 
  st_drop_geometry()

###

plan(callr, workers = 6)
options(future.globals.maxSize = 1.6e+10)

groups <- split(1:10612, ceiling(seq_along(1:10612)/25))

###

for (x in 1:length(groups)) {
  int_vector <- unlist(groups[x])
  
  ok_table <- future_map(int_vector, ~router(.x, destination = ok_dest)) %>% 
  bind_rows()
  
  ok_path <- paste0('outputs/network_components/routes/okanagan/okanagan_route_', x, '.rds')
  write_rds(ok_table, ok_path)
  print(paste('Completed:', x))
  gc()
}

```

# Summarize route data

```{r}

okanagan_files <- list.files('outputs/network_components/routes_compiled_spatial/', 
                            pattern = '*okanagan',
                            full.names = TRUE,
                            recursive = TRUE) 

template <- tibble(index = character(),
                            destination_id = numeric(),
                            origin_id = numeric(),
                            node_position = numeric(),
                            x = numeric(),
                            y = numeric(),
                    metro_id = numeric(),
                    metro_name = character(),
                    prov_state_name = character(),
                    country = character(),
                    buffer_dist = numeric()
                    )

for (i in 1:length(okanagan_files)) {
  ival = eval(i)
  intermediate <- okanagan_files[ival]
  intermediate <- read_rds(intermediate)
  intermediate <- intermediate %>% filter(destination_id == 1)
  template <- bind_rows(template, intermediate)
  print(paste('Completed:', ival))
}

test_fraction <- template %>% filter(destination_id == 1)


test_grouped <- test_fraction  %>% group_by(origin_id) %>% 
  arrange(desc(node_position)) %>% 
  rowid_to_column('route_index') %>% 
  mutate(origin_id_fixed = cur_group_id()) %>% 
  ungroup() %>% 
group_by(origin_id_fixed)

test_grouped$origin_id_fixed <- as.character(test_grouped$origin_id_fixed)

test_grouped <- test_grouped %>% filter(origin_id_fixed < 20)

test_group_line_1 <- summarize(test_grouped, route_index = min(route_index))

test_group_line_2 <- summarize(test_grouped, route_index = max(route_index))

test_group_line <- bind_rows(test_group_line_1, test_group_line_2) %>% 
  group_by(origin_id_fixed)

filtered <- test_grouped %>% ungroup() %>% group_by(origin_id_fixed, metro_name) %>% 
  summarize()

test_group_line <- left_join(test_group_line, filtered)

test_grouped <- test_grouped %>% drop_na()

slice_example <- ggplot() +  
  geom_line(test_group_line, mapping = aes(x = route_index, y = origin_id_fixed, 
                                           colour = metro_name),
            alpha = 0.7, linewidth = 0.75) +
 paletteer::scale_color_paletteer_d("fishualize::Oncorhynchus_gorbuscha") +
    guides(colour = guide_legend('Origin city')) +
  new_scale_color() +
  geom_point(test_grouped, mapping = aes(x = route_index, y = origin_id_fixed,
                                        colour = buffer_dist),
             position = position_dodge(0.5)) +
    ylab('Origin ID') +
  xlab('Route index') +
  scale_colour_paletteer_c("ggthemes::Sunset-Sunrise Diverging",
                na.value = 'grey80',
                direction = -1) +
  guides(colour = guide_legend('Distance \n to zebra mussel \n infestation')) +
  theme_classic(base_family = extrafont::choose_font('Source Code Pro'), 
                base_size = 12) 

slice_example

ggsave('figures/meeting_figures/feb_27_2024/slice_example.png', 
       width = 9, 
       height = 12, 
       units = 'in')


```


# route data part 2

```{r}


okanagan_files <- list.files('outputs/network_components/routes_compiled_spatial/', 
                            pattern = '*okanagan',
                            full.names = TRUE,
                            recursive = TRUE) 

outer_template <- tibble(destination_id = numeric(),
                         origin_id = numeric(),
                         node_position = numeric(),
                         metro_id = numeric(),
                         metro_name = character(),
                         prov_state_name = character(),
                         country = character(),
                         buffer_dist = numeric(),
                         non_na_count = numeric(),
                         na_count = numeric(),
                         sum_obs = numeric(),
                         mean_buff_dist = numeric(),
                         sd_buffer_dist = numeric(),
                         scaled_exposure = numeric(),
                         rough_distance_km = numeric())

for (i in 1:length(okanagan_files)) {
  
  ival = eval(i)
  
  origin_file <- read_rds(okanagan_files[ival]) %>%
    select(-x, -y, -index)
  
  de_duplicated <- origin_file %>% 
  filter(!duplicated(paste0(pmax(destination_id, origin_id), 
                            pmin(destination_id, origin_id)))) %>% 
    select(-buffer_dist)
  
  observations <- origin_file %>% group_by(destination_id, origin_id) %>% 
  summarise(sum_obs = n())
  
  dat_frame <- origin_file %>% drop_na() %>% 
    group_by(destination_id, origin_id) %>% 
    count(buffer_dist) %>% 
    summarize(non_na_count = sum(n))
  
  na_frame <- origin_file %>% 
    group_by(destination_id, origin_id) %>%
    count(buffer_dist) %>% 
    filter(is.na(buffer_dist)) %>% 
    select(-buffer_dist) %>% 
    rename(na_count = n)
  
  sum_stats <- origin_file %>% 
    group_by(destination_id, origin_id) %>%
    drop_na() %>% 
    summarize(mean_buff_dist = mean(buffer_dist), sd_buffer_dist = sd(buffer_dist))
  
  template <- left_join(de_duplicated, dat_frame, by = c('destination_id', 'origin_id'))
  
  template <- left_join(template, na_frame, by = c('destination_id', 'origin_id'),
                      keep = FALSE)
  template <- left_join(template, observations, by = c('destination_id', 'origin_id'),
                      keep = FALSE)
  template <- left_join(template, sum_stats, by = c('destination_id', 'origin_id'),
                      keep = FALSE)
  
  data_sheet <- template %>% mutate(scaled_exposure =  non_na_count/sum_obs,
                                  rough_distance_km = sum_obs*2)
  
  outer_template <- bind_rows(outer_template, data_sheet)
  
  print(paste('Completed:', ival))
}

write_rds(outer_template, 'outputs/analysis/okanagan_route_data.rds')


demonstration_filter <- outer_template %>% filter(destination_id == 14)

origin_collection <- read_sf('outputs/network_components/origin_matched.gpkg')

demonstration_spatial <- right_join(origin_collection, demonstration_filter, 
                                    by = 'origin_id', keep = FALSE)

write_sf(demonstration_spatial, 'outputs/analysis/okanagan_origins.gpkg')


metro_centroids <- st_centroid(north_america_metros)%>% 
  mutate(x = sf::st_coordinates(.)[,1],
         y = sf::st_coordinates(.)[,2]) %>% 
  st_drop_geometry()
  


demonstration_spatial_cities <- demonstration_spatial %>% group_by(metro_name.x) %>% 
  summarize(mean_exposure = mean(scaled_exposure),
            sd_exposure = sd(scaled_exposure)) %>% 
  replace_na(list(mean_exposure = 0, sd_exposure = 0)) %>% 
  rename(metro_name = metro_name.x) %>% 
  left_join(metro_centroids, by = 'metro_name')

  

city_list <- ggplot() + geom_point(demonstration_spatial_cities, mapping = aes(x = metro_id, 
                                                        y = mean_exposure, 
                                                        colour = mean_exposure)) +
  paletteer::scale_color_paletteer_c("ggthemes::Orange-Gold", direction = -1) +
      ylab('Mean exposure') +
  xlab('Cities') +
  guides(colour = guide_legend('Exposure index')) +
  theme_classic(base_family = extrafont::choose_font('Source Code Pro'), 
                base_size = 12) 

city_list


west_east <- ggplot() + geom_point(demonstration_spatial_cities, mapping = aes(x = x, 
                                                        y = mean_exposure, 
                                                        colour = mean_exposure)) +
paletteer::scale_color_paletteer_c("ggthemes::Orange-Gold", direction = -1) +
  ylab('Mean exposure') +
  xlab('West-East') +
  guides(colour = guide_legend('Exposure index')) +
  theme_classic(base_family = extrafont::choose_font('Source Code Pro'), 
                base_size = 12) 

west_east

south_north <- ggplot() + geom_point(demonstration_spatial_cities, mapping = aes(x = y, 
                                                        y = mean_exposure, 
                                                        colour = mean_exposure)) +
  paletteer::scale_color_paletteer_c("ggthemes::Orange-Gold", direction = -1) +
  ylab('Mean exposure') +
  xlab('South-North') +
  guides(colour = guide_legend('Exposure index')) +
  theme_classic(base_family = extrafont::choose_font('Source Code Pro'), 
                base_size = 12) 

south_north


```



# Raster gen  

```{r}


okanagan_files <- list.files('outputs/network_components/routes_compiled_spatial/', 
                            pattern = '*okanagan',
                            full.names = TRUE,
                            recursive = TRUE) 

land <- read_sf('processed/shared_north_america_coastline.gpkg') %>% 
  st_transform(crs = 'ESRI:102008') %>% vect() %>% 
  rast(res = 2000)






for (i in 1:length(okanagan_files)) {
  ival = eval(i)
  intermediate <- okanagan_files[ival]
  intermediate <- read_rds(intermediate)
  intermediate <- intermediate %>% select(x, y)
  intermediate <- intermediate %>%
    st_as_sf(coords = c('x', 'y'), 
             crs = st_crs('ESRI:102008'))
  pathname <- paste0('side_analyses/raster_generation/point_collection/pt_subset_', ival, '.gpkg')
  write_sf(intermediate, pathname)
  print(paste('Completed:', ival))
}


overlay_grid <- read_sf('processed/2km_hexagon_grid.gpkg')%>% 
  rowid_to_column('hex_id')

overlay_grid <- overlay_grid %>% mutate(x = sf::st_coordinates(.)[,1],
         y = sf::st_coordinates(.)[,2])

hex_id_template <- overlay_grid %>% st_drop_geometry()

write_rds(hex_id_template, 'side_analyses/raster_generation/auxiliary_files/hex_id_template.rds')


okanagan_spatial_files <- list.files('side_analyses/raster_generation/point_collection/', 
                            pattern = '*.gpkg',
                            full.names = TRUE,
                            recursive = TRUE) 

for (i in 1:length(okanagan_spatial_files)) {
  ival = eval(i)
  intermediate <- read_sf(okanagan_spatial_files[ival]) %>% 
    mutate(count_number = 1)
  count_name <- paste0('count_', ival)
  suppressWarnings(overlay_hex <- st_intersection(intermediate, overlay_grid) %>% 
    group_by(hex_id) %>% 
    count() %>% 
    st_drop_geometry() %>% 
    rename(!!count_name := n)) 
  pathname <- paste0('side_analyses/raster_generation/hex_joined/hex_joined_', ival, '.rds')
  write_rds(overlay_hex, pathname)
  print(paste('Completed:', ival))
  rm(intermediate)
  gc()
}





hex_files <- list.files('side_analyses/raster_generation/hex_joined/', 
                            pattern = '*.rds',
                            full.names = TRUE,
                            recursive = TRUE) 

hex_id_template <- read_rds('side_analyses/raster_generation/auxiliary_files/hex_id_template.rds')


start <- c(1, 25, 50, 75, 100, 125, 150, 175, 200, 
           225, 250, 275, 300, 325, 350, 375, 400)

end <- c(25, 50, 75, 100, 125, 150, 175, 200, 
         225, 250, 275, 300, 325, 350, 375, 400, 422)

for (j in 1:17) {
  j_val <- eval(j)
  start_eval <- start[j_val]
  end_eval <- end[j_val]
  file_filtered <- hex_files[start_eval:end_eval]
  print(paste('Starting group:', j_val))
  
  hex_id_template <- read_rds('side_analyses/raster_generation/auxiliary_files/hex_id_template.rds')
  
  for (i in 1:length(file_filtered)) {
    ival = eval(i)
    intermediate <- read_rds(file_filtered[ival])
    hex_id_template <- left_join(hex_id_template, intermediate, by = 'hex_id')
    print(paste('Completed:', ival))
}
  filename <- paste0('side_analyses/raster_generation/hex_collated/collated_', j_val, '.rds')
  
  write_rds(hex_id_template, filename)
  
  print(paste('Completed group:', j_val))
  
  rm(hex_id_template)
  gc()
  
}


collate_list <- list.files('side_analyses/raster_generation/hex_collated/', 
                            pattern = '*.rds',
                            full.names = TRUE,
                            recursive = TRUE) 


test <- read_rds(collate_list[1]) %>% 
  replace(is.na(.), 0) %>% transform(hex_id = as.character(hex_id)) %>% 
  mutate(total_count = rowSums(across(where(is.numeric)), na.rm=TRUE)) %>% 
  select(hex_id, total_count)

```


```{r}

# Exposure metrics:

# Cumulative exposures - # of exposure events in each bracket/number of points

# Average exposure distance - of exposure events, what is average distance



```
