# Zebra mussel network modeling script

# Contact information

Contact: Steven Brownlee
Email: steven.fr.brownlee@gmail.com
Date last revised: June 24 2024

# Session info

R version 4.4.1 (2024-06-14)
Platform: x86_64-redhat-linux-gnu
Running under: Nobara Linux 40 (GNOME Edition)

Matrix products: default
BLAS/LAPACK: FlexiBLAS OPENBLAS-OPENMP;  LAPACK version 3.11.0

locale:
 [1] LC_CTYPE=en_CA.UTF-8      LC_NUMERIC=C              LC_TIME=en_CA.utf8       
 [4] LC_COLLATE=en_CA.UTF-8    LC_MONETARY=en_CA.utf8    LC_MESSAGES=en_CA.UTF-8  
 [7] LC_PAPER=en_CA.utf8       LC_NAME=C                 LC_ADDRESS=C             
[10] LC_TELEPHONE=C            LC_MEASUREMENT=en_CA.utf8 LC_IDENTIFICATION=C      

time zone: America/Vancouver
tzcode source: system (glibc)

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

loaded via a namespace (and not attached):
 [1] digest_0.6.36     tidyr_1.3.1       utf8_1.2.4        R6_2.5.1         
 [5] fastmap_1.2.0     tidyselect_1.2.1  xfun_0.46         magrittr_2.0.3   
 [9] glue_1.7.0        tibble_3.2.1      knitr_1.48        pkgconfig_2.0.3  
[13] htmltools_0.5.8.1 rmarkdown_2.27    dplyr_1.1.4       generics_0.1.3   
[17] lifecycle_1.0.4   cli_3.6.3         fansi_1.0.6       vctrs_0.6.5      
[21] compiler_4.4.1    purrr_1.0.2       rstudioapi_0.16.0 tools_4.4.1      
[25] pillar_1.9.0      evaluate_0.24.0   yaml_2.3.10       rlang_1.1.4   

# Set up data directory

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/home/steven/Documents/workspace/thesis/ch2/gis_data/')
```

# R library setup

```{r}
library(sf)
library(tidyverse)
library(rcartocolor)
library(sfnetworks)
library(qgisprocess)
library(future)
library(future.callr)
library(purrr)
library(furrr)
library(progressr)

```

# Import shapefiles

```{r}
zm_occurrences <- read_sf('processed/dreissenid_occurrences.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

us_metro <- read_sf('processed/us_metro_areas.gpkg')%>% 
  st_transform(crs = 'ESRI:102008')

canada_metro <- read_sf('processed/canada_metro_areas.gpkg')%>% 
  st_transform(crs = 'ESRI:102008')

land <- read_sf('processed/shared_north_america_coastline.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

lakes <- read_sf('processed/cec_lakes.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

usdot_road_network <- read_sf('processed/usdot_road_network.gpkg')%>% 
  st_transform(crs = 'ESRI:102008')

qgis_bbox <- read_sf('processed/qgis_bbox.gpkg') %>% 
  st_transform(crs = 'ESRI:102008')

```


# Pre-processing

```{r}
# Create buffers around ZM occurrences

zm_10 <- st_buffer(zm_occurrences, dist = 10000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '10 km')

zm_20 <- st_buffer(zm_occurrences, dist = 20000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '20 km')

zm_30 <- st_buffer(zm_occurrences, dist = 30000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '30 km')

zm_40 <- st_buffer(zm_occurrences, dist = 40000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '40 km')

zm_50 <- st_buffer(zm_occurrences, dist = 50000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '50 km')

zm_buffer_collected <- rbind(zm_50, zm_40, zm_30, zm_20, zm_10)

rm(zm_50, zm_40, zm_30, zm_20, zm_10)

zm_buffer_collected$buffer_dist <- factor(zm_buffer_collected$buffer_dist, levels = c('50 km', '40 km', '30 km', '20 km', '10 km'))

zm_buffer_collected <- st_transform(zm_buffer_collected, crs = 4326)

write_sf(zm_buffer_collected, 'processed/zm_buffer.gpkg')

# Merge US and Canada metro areas

us_metro <- us_metro %>% select(NAME) %>% rename(metro_area = NAME)

canada_metro <- canada_metro %>% select(CMANAME) %>% rename(metro_area = CMANAME)

north_america_metro <- rbind(us_metro, canada_metro)

# Create node network for USDOT roads
```


```{r}

usdot_road_network <- usdot_road_network %>% rownames_to_column('seg_id')


plan(callr, workers = 6)
options(future.globals.maxSize= 891289600)

node_generator <- function(sublist) {

for (x in sublist) {
  usdot_template <- tibble(x_coord = numeric(),
                         y_coord = numeric(),
                         country = character(),
                         usdot_id = numeric(),
                         jurisdiction = character(),
                         road_number = character(),
                         road_name = character(),
                         surface_type = character(),
                         lanes = numeric(),
                         speed_limit = numeric(),
                         class = numeric())
  
segment_filtered <- usdot_road_network %>% filter(seg_id == x)
length_extract = segment_filtered$LENGTH

if (length_extract <= 1) {
  sample_size = 5
} else if (1 < length_extract & length_extract <= 5 ) {
  sample_size = 20
} else if (5 < length_extract & length_extract <= 20) {
  sample_size = 100 
} else if (20 < length_extract & length_extract <= 50) {
  sample_size = 400 
} else if (50 < length_extract & length_extract <= 100) {
  sample_size = 700
} else if (100 < length_extract & length_extract <= 400) {
  sample_size = 2500
} else {sample_size = 2000} 

usdot_nodes <- st_sample(segment_filtered, size = sample_size, type = 'regular') %>% 
  st_cast("POINT") %>% 
  st_as_sf() %>% 
  mutate(x_coord = sf::st_coordinates(.)[,1],
         y_coord = sf::st_coordinates(.)[,2],
         usdot_id = segment_filtered$ID,
         country = segment_filtered$COUNTRY,
         jurisdiction = segment_filtered$JURISNAME,
         road_number = segment_filtered$ROADNUM,
         road_name = segment_filtered$ROADNAME,
         surface_type = segment_filtered$SURFACE,
         lanes = segment_filtered$LANES,
         speed_limit = segment_filtered$SPEEDLIM,
         class = segment_filtered$CLASS) %>% 
  st_drop_geometry()

usdot_nodes$x_coord <- as.numeric(usdot_nodes$x_coord)
usdot_nodes$y_coord <- as.numeric(usdot_nodes$y_coord)

usdot_template <- rbind(usdot_template, usdot_nodes) 

}
usdot_filled <- usdot_template
return(usdot_filled)
}

usdot_nodes <- future_map(1:716638, ~node_generator(.x))

###

usdot_nodes_2 <- bind_rows(usdot_nodes)

write_rds(usdot_nodes_2, 'outputs/node_folder/usdot_combined_2.rds')

rm(usdot_nodes_2, usdot_nodes)

usdot_combined <- read_rds('outputs/node_folder/usdot_combined_2.rds') 

usdot_combined <- usdot_combined %>% rowid_to_column('node_id')


inside_list = c(1, 500000,
                1000000,
                2000000,
                3000000,
                4000000,
                5000000,
                6000000,
                7000000,
                8000000,
                9000000,
                10000000,
                11000000)

outside_list= c(500000,
                1000000,
                2000000,
                3000000,
                4000000,
                5000000,
                6000000,
                7000000,
                8000000,
                9000000,
                10000000,
                11000000,
                11544435)

for (i in 1:13) {
  inside_index <- inside_list[i]
  outside_index <- outside_list[i]
  usdot_subset <- usdot_combined %>% 
    filter(node_id >= inside_index & node_id < outside_index)
   usdot_subset <- usdot_subset %>% 
     drop_na() %>% 
  st_as_sf(coords = c('x_coord', 'y_coord'), crs = st_crs('ESRI:102008'))
  write_sf(usdot_subset, paste0('outputs/node_folder/node_pieces/node_subset_', i, '.gpkg'))
  print(paste('Completed:', i))
}

###
```


```{r}
#overlay_grid <- st_make_grid(qgis_bbox, cellsize = 2000,
#                             what = 'polygons', square = FALSE) %>% st_as_sf() %>% 
#  st_transform(crs = 'ESRI:102008')

#write_sf(overlay_grid, 'processed/2km_hexagon_grid.gpkg')

overlay_grid <- read_sf('processed/2km_hexagon_grid.gpkg')

overlay_grid <- overlay_grid %>% rowid_to_column('hex_id')

overlay_grid$hex_group <- as.numeric(cut_number(overlay_grid$hex_id, 1000))

for (x in 1:1000) {
  overlay_subset <- overlay_grid %>% filter(hex_group == x)
  write_sf(overlay_subset, paste0('outputs/overlay_grid_segments/hex_group_', x, '.gpkg'))
  print(paste('Completed:', x))
}


hex_list <- list.files('outputs/overlay_grid_segments/', 
                            pattern = '*.gpkg',
                            full.names = TRUE,
                            recursive = TRUE)

node_list <- list.files('outputs/node_folder/node_pieces/', 
                            pattern = '*.gpkg',
                            full.names = TRUE,
                            recursive = TRUE)


table_template <- tibble(hexid = numeric(),
                         node_id = numeric(),
                         usdot_id = numeric(),
                         country = numeric(),
                         jurisdiction = character(),
                         road_number = character(),
                         road_name = character(),
                         surface_type = character(),
                         lanes = numeric())

joiner <- function(sequence) {
for (x in sequence) {
  hex_segment  <- read_sf(hex_list[x]) %>% select(-hex_group)
  for (i in 1:13) {
    interior_nodes  <- read_sf(node_list[i])
    hex_interior <- st_join(hex_segment, interior_nodes) %>% st_drop_geometry()
    table_template <- rbind(table_template, hex_interior)
    print(paste('Interior done:', i))
}
  table_export <- table_template
  write_rds(table_export, paste0('outputs/overlay_grid_joined/overlay_grid_joined_', x, '.rds'))
  table_template <- tibble(hexid = numeric(),
                         node_id = numeric(),
                         usdot_id = numeric(),
                         country = numeric(),
                         jurisdiction = character(),
                         road_number = character(),
                         road_name = character(),
                         surface_type = character(),
                         lanes = numeric())
  return(table_export)
}
}

plan(callr, workers = 6)
options(future.globals.maxSize= 891289600)


joined <- future_map(1:1000, ~joiner(.x))

joined <- bind_rows(joined)

write_rds(joined, 'outputs/overlay_grid_joined/overlay_tables/joined_table.rds')

joined_table <- drop_na(joined)

write_rds(joined_table, 'outputs/overlay_grid_joined/overlay_tables/joined_table_na_dropped.rds')

joined_table <- read_rds('outputs/overlay_grid_joined/overlay_tables/joined_table_na_dropped.rds')

joined_table_summarized <- joined_table %>% select(hex_id, country, jurisdiction,
                                                   surface_type, surface_type,
                                                   lanes, speed_limit, class)

joined_table_summarized$surface_type <- case_match(joined_table_summarized$surface_type, 
                                            'Unpaved' ~ '1', 
                                            'Paved' ~ '2')

joined_table_summarized$surface_type <- as.numeric(joined_table_summarized$surface_type)

joined_table_summarized <- joined_table %>% 
  group_by(hex_id) %>% 
  summarize(surface_type = max(surface_type),
            lanes = max(lanes),
            speed_limit = max(speed_limit))

hex_final <- left_join(overlay_grid, joined_table_summarized, by = 'hex_id') %>% 
  drop_na()


write_sf(hex_final, 'outputs/network_components/hex_grid_final.gpkg')

```

```{r}
hexagon_grid <- read_sf('outputs/network_components/hex_grid_final.gpkg') %>% 
  rowid_to_column('hex_rowid')

hex_centroids <- hexagon_grid %>% st_centroid() %>%  
  mutate(x = sf::st_coordinates(.)[,1],
         y = sf::st_coordinates(.)[,2]) 

hex_centroids_aspatial <- hex_centroids %>% st_drop_geometry()

###

row_indices <- split(hex_centroids$hex_rowid, cut(hex_centroids$hex_rowid, breaks = 160))


###

edge_generator <- function(row_subset) {
  
  distance_array_template <- tibble(hex_rowid_1 = numeric(),
                                    x_1 = numeric(),
                                    y_1 = numeric(),
                                    hex_rowid_2 = numeric(),
                                    x_2 = numeric(),
                                    y_2 = numeric(),
                                    comb_hex_rowid = character())
  
  for (x in row_subset) {
    xval = eval(x)
    coord_1 <- hex_centroids %>% filter(hex_rowid == xval)
    matches <- as_tibble(as.data.frame(st_is_within_distance(coord_1, hex_centroids, 
                                                      dist = 2500))) %>% 
      select(col.id) %>% deframe()
    
    coord_matches <- hex_centroids_aspatial %>% 
      filter(hex_rowid %in% matches) %>% 
      select(hex_rowid, x, y) %>% 
      mutate(hex_rowid_1 = coord_1$hex_id,
             x_1 = coord_1$x,
             y_1 = coord_1$y) %>% 
      rename(hex_rowid_2 = hex_rowid,
             x_2 = x,
             y_2 = y) %>% 
      mutate(comb_id = paste0(as.character(hex_rowid_1), '_', as.character(hex_rowid_2)))
   
     distance_array_template <- rbind(distance_array_template, coord_matches)
  }
  return(distance_array_template)
}

# Florilegium: 

plan(callr, workers = 6)
options(future.globals.maxSize = 891289600)  

for  (i in 1:40) {
  ival = eval(i)
  edge_collection <- future_map(row_indices[ival], ~edge_generator(.x)) %>% bind_rows
  write_rds(edge_collection, paste0('outputs/network_components/edge_folder/edge_collection_',
                                   ival, '.rds'))
  print(paste('Completed:', ival))
}


# Advesperascit:

plan(callr, workers = 12)
options(future.globals.maxSize = 891289600)

for  (i in 40:80) {
  ival = eval(i)
  edge_collection <- future_map(row_indices[ival], ~edge_generator(.x)) %>% bind_rows
  write_rds(edge_collection, paste0('outputs/network_components/edge_folder/edge_collection_',
                                   ival, '.rds'))
  print(paste('Completed:', ival))
}


# Cognoscere

plan(callr, workers = 20)
options(future.globals.maxSize = 891289600)

for  (i in 80:120) {
  ival = eval(i)
  edge_collection <- future_map(row_indices[ival], ~edge_generator(.x)) %>% bind_rows
  write_rds(edge_collection, paste0('/home/steven/Downloads/edge_collection_',
                                   ival, '.rds'))
  print(paste('Completed:', ival))
}
  
####

edge_compilation <- list.files('outputs/network_components/edge_folder/', 
                            pattern = '*.rds',
                            full.names = TRUE,
                            recursive = TRUE)

edge_compilation <- lapply(edge_compilation, readRDS)


edge_compilation <- edge_compilation %>% 
mutate(geometry = paste0('LINESTRING (', x_1, ' ', y_1, ', ', x_2, ' ', y_2, ')' ))

edge_compilation <- edge_compilation %>% st_as_sf(wkt = 'geometry', crs = 'ESRI:102008')

write_sf(edge_compilation, '/home/steven/Downloads/edge_test.gpkg')


test <- readRDS('/home/steven/Downloads/edge_collection_1.rds') %>% 
mutate(geometry = paste0('LINESTRING (', x_1, ' ', y_1, ', ', x_2, ' ', y_2, ')' )) %>% 
  st_as_sf(wkt = 'geometry', crs = 'ESRI:102008')

write_sf(test, '/home/steven/Downloads/edge_revised.gpkg')


```

