# Zebra mussel network modeling script

# Contact information

Contact: Steven Brownlee
Email: steven.fr.brownlee@gmail.com
Date last revised: June 24 2024

# Session info

R version 4.4.1 (2024-06-14)
Platform: x86_64-redhat-linux-gnu
Running under: Nobara Linux 40 (GNOME Edition)

Matrix products: default
BLAS/LAPACK: FlexiBLAS OPENBLAS-OPENMP;  LAPACK version 3.11.0

locale:
 [1] LC_CTYPE=en_CA.UTF-8      LC_NUMERIC=C              LC_TIME=en_CA.utf8       
 [4] LC_COLLATE=en_CA.UTF-8    LC_MONETARY=en_CA.utf8    LC_MESSAGES=en_CA.UTF-8  
 [7] LC_PAPER=en_CA.utf8       LC_NAME=C                 LC_ADDRESS=C             
[10] LC_TELEPHONE=C            LC_MEASUREMENT=en_CA.utf8 LC_IDENTIFICATION=C      

time zone: America/Vancouver
tzcode source: system (glibc)

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

loaded via a namespace (and not attached):
 [1] digest_0.6.36     tidyr_1.3.1       utf8_1.2.4        R6_2.5.1         
 [5] fastmap_1.2.0     tidyselect_1.2.1  xfun_0.46         magrittr_2.0.3   
 [9] glue_1.7.0        tibble_3.2.1      knitr_1.48        pkgconfig_2.0.3  
[13] htmltools_0.5.8.1 rmarkdown_2.27    dplyr_1.1.4       generics_0.1.3   
[17] lifecycle_1.0.4   cli_3.6.3         fansi_1.0.6       vctrs_0.6.5      
[21] compiler_4.4.1    purrr_1.0.2       rstudioapi_0.16.0 tools_4.4.1      
[25] pillar_1.9.0      evaluate_0.24.0   yaml_2.3.10       rlang_1.1.4   

# Set up data directory

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/home/steven/Documents/common/thesis/ch2/gis_data/')
```

# R library setup

```{r}
library(sf)
library(tidyverse)
library(rcartocolor)
library(sfnetworks)

```

# Import shapefiles

```{r}
zm_occurrences <- read_sf('data_directory/usgs_nas_dreissenid_occurrences.shp') %>% st_transform(crs = 4326)

write_sf(zm_occurrences, 'data_directory/usgs_nas_dreissenid_occurrences.shp')

north_america_bbox <- read_sf('data_directory/qgis_north_america_bbox.shp') %>% st_transform(crs = 4326)

write_sf(north_america_bbox, 'data_directory/qgis_north_america_bbox.shp')


us_major_cities_centroid <- read_sf('data_directory/cec_populated_places.shp') %>% st_transform(crs = 4326)

write_sf(us_major_cities_centroid, 'data_directory/cec_populated_places.shp')


land <- read_sf('data_directory/qgis_cec_coastline.shp') %>% st_transform(crs = 4326)

write_sf(land, 'data_directory/qgis_cec_coastline.shp')


large_lakes <- read_sf('data_directory/cec_lakes_shapefile.shp') %>% st_transform(crs = 4326)

write_sf(large_lakes, 'data_directory/cec_lakes_shapefile.shp')


usdot_road_network <- read_sf('downloads/us_dot_road_network/NTAD_North_American_Roads_2691167086962748384.gpkg') %>% st_transform(crs = 4326)


write_sf(usdot_road_network, 'data_directory/usdot_road_network.shp')


```


# Pre-processing

```{r}

zm_established <- zm_occurrences %>% filter(Status == 'established') %>% 
  filter(Scientific == 'Dreissena polymorpha')

zm_10 <- st_buffer(zm_established, dist = 10000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '10 km')

zm_20 <- st_buffer(zm_established, dist = 20000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '20 km')

zm_30 <- st_buffer(zm_established, dist = 30000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '30 km')

zm_40 <- st_buffer(zm_established, dist = 40000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '40 km')

zm_50 <- st_buffer(zm_established, dist = 50000) %>% st_union() %>% st_as_sf() %>% 
  mutate(buffer_dist = '50 km')
 
 
zm_buffer_collected <- rbind(zm_50, zm_40, zm_30, zm_20, zm_10)

rm(zm_50, zm_40, zm_30, zm_20, zm_10)

zm_buffer_collected$buffer_dist <- factor(zm_buffer_collected$buffer_dist, levels = c('50 km', '40 km', '30 km', '20 km', '10 km'))


```



```{r}

overlay_grid <- st_make_grid(north_america_bbox, cellsize = 2000, what = 'polygons',
                             square = FALSE) %>% st_as_sf()

write_sf(overlay_grid, 'scratch/2km_hexagon_grid.gpkg')

overlay_selection <- qgis_run_algorithm(
  "native:extractbylocation",
  INPUT = 'scratch/10km_hexagon_grid.gpkg',
  PREDICATE = 6,
  INTERSECT = 'data_directory/qgis_cec_coastline.shp',
  OUTPUT = 'scratch/10km_hexagon_selection.gpkg'
)

overlay_selection <- read_sf('scratch/2km_hexagon_selection.gpkg')

overlay_selection <- overlay_selection %>% rowid_to_column('uid') %>% select(uid)

write_sf(overlay_selection, 'data_directory/2km_hexagon_selection.gpkg', driver = 'GPKG')

```

# 

```{r}

overlay_selection <- read_sf('data_directory/2km_hexagon_selection.gpkg') %>% st_transform(crs = 4326)

write_sf(overlay_selection,'data_directory/2km_hexagon_selection.gpkg')

usdot_road_network <- usdot_road_network %>% filter(SURFACE == 'Unpaved' | SURFACE == 'Paved')

overlay_usdot <- st_join(overlay_selection, usdot_road_network, join = st_intersects)

test <- overlay_usdot %>% drop_na()

write_sf(test,'scratch/2km_hexagon_joined_test.gpkg')

```



```{r}

zm_bbox <- st_bbox(north_america_bbox)


buffer_plot <- ggplot() + geom_sf(data = land, fill = 'gray70', colour = 'gray15') +
  geom_sf(data = large_rivers, colour = 'gray60') +
  geom_sf(data = small_rivers, colour = 'gray60', alpha = 0.3) +
    geom_sf(data = large_lakes, fill = 'gray50', colour = 'gray15') +
  geom_sf(data = small_lakes, fill = 'gray50', colour = 'gray15') +
  geom_sf(data = zm_buffer_collected_test, aes(fill = buffer_dist)) +
  scale_fill_carto_d(palette = 'TealGrn', direction = -1, name = 'Buffer distance \n in km') +
  theme_dark(base_family = extrafont::choose_font('Cantarell Light'), 
               base_size = 12) +
      coord_sf(xlim = c(zm_bbox[[1]], zm_bbox[[3]]),
             ylim = c(zm_bbox[[2]], zm_bbox[[4]]), 
             expand = FALSE) 



```

```{r}

test <- oe_read(file_path = '/home/steven/Documents/common/thesis/ch2/gis_data/downloads/open_street_map/north-america-latest.osm.pbf', layer = 'lines')

```


Run:

usethis::edit_r_environ()

OSMEXT_DOWNLOAD_DIRECTORY=/home/steven/Documents/common/thesis/ch2/gis_data/downloads/open_street_map_r

```{r}


overlay_selection <- read_sf('data_directory/2km_hexagon_selection.gpkg')

test_filter <- overlay_selection %>% filter(uid == 496538)

            
              
      

canada <- oe_vectortranslate('downloads/open_street_map_r/geofabrik_canada-latest.osm.pbf')

canada_gpkg <- oe_read('downloads/open_street_map_r/geofabrik_canada-latest.gpkg')

canada_gpkg_test <- canada_gpkg %>% 
  filter(highway == "motorway" | highway == "tertiary" | highway == "residential" | 
           highway == "primary" | highway == "unclassified" | highway ==  "road" |
           highway == "trunk" | highway == "secondary" | highway == "motorway_link")

write_sf(canada_gpk_test, 'data_directory/')

bc <- oe_get('British Columbia', max_file_size = 4e+10, 
                 query = "SELECT * FROM 'lines' WHERE highway = 'motorway') 
                 OR highway = 'trunk' OR highway = 'primary' OR highway = 'secondary' OR highway = 'tertiary' OR highway = 'unclassified'  OR highway = 'residential' OR highway = 'motorway_link' OR highway = 'trunk_link' OR highway = 'primary_link' OR highway = 'secondary_link' OR highway = 'tertiary_link' OR highway = 'service' OR highway = 'track'")

```

```{r}

```

